---
title: "Metabolite screen Microbacteria"
subtitle: "Thoenen_et_al_Bacteria_BX_metabolization"
author: "Lisa Thönen"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_caption: yes
    toc: true
    toc_float: true
    theme: cerulean
    code_folding: show
editor_options: 
  chunk_output_type: console
---

Here we plot the results of the metabolisation assay in liquid cultures of Microbacteria from the maize root bacteria (Thoenen et al. 2023) and the AtSphere collection (Bai et al. 2015). We tested MBOA and DIMBOA-Glc. Additionally we also report growth data of all Microbacteria in MBOA-containing TSB medium to test them for MBOA tolerance and in MBOA-containing minimal medium to test them for using MBOA as sole carbon source for growth.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
rm(list = ls())
```

```{r,include=FALSE}
# install.packages("dplyr")
# install.packages("magrittr") # for the %>% function
# install.packages("ggplot2") # plots
# install.packages("readxl") # to read excels
# install.packages("stringr")
# install.packages("MESS") # for auc function
# install.packages("purrr")
# install.packages("multcomp")
# install.packages("emmeans")
# install.packages("ggpubr")
# install.packages("forcats")
# install.packages("readr")
# install.packages("ggpmisc")
# install.packages("pander")
# install.packages("rstatix")
# install.packages("ggtree")
# install.packages("tibble")
# install.packages("aplot")
# install.packages("knitr")
# install.packages("ggnewscale")
# install.packages("svglite")
```

```{r, warning=F, echo=F, message=F}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("ggtree")
```

```{r,include=FALSE}
library("dplyr")
library("magrittr")
library("ggplot2")
library("readxl")
library("tidyr")
library("readr")
library("stringr")
library("MESS") 
library("purrr")
library("ggpubr")
library("forcats")
library("readr")
library("ggpmisc")
library("pander")
library("ggtree")
library("tibble")
library("aplot")
library("knitr")
library("ggnewscale")
library("svglite")
```

```{r functions, echo=F, message=F}
give.n <- function(x){
  return(c(y = median(x)*1.05, label = length(x))) 
  # experiment with the multiplier to find the perfect position
}
```

# MBOA metabolisation 

```{r, warning=F, echo=F, message=F}
# Metadata
metadata <- read_excel("Input/210513_experimental_design_Microbacteria_MRB_At_metabolites.xlsx", sheet = 6)
# Metabolite data
met <- read_excel("Input/20211007_Microbacteria_metabolites.xlsx", sheet = 2)
# Molecular weights
mol <- read_excel("Input/Molecular_weight_BXDs.xlsx")

mol %<>% mutate(Compound = gsub("MHPA", "HMPAA", Compound)) %>% 
       mutate(Compound = gsub("MAPH", "MAP",  Compound)) 
# more info strains
database <- read_excel("Input/210513_experimental_design_Microbacteria_MRB_At_metabolites.xlsx", sheet = 2)
```

```{r, warning=F, echo=F, message=F}
met %<>% dplyr::rename(Std_conc = `Std. Conc`, ng_mL_ML = `ng/mL`) %>% 
       mutate(Compound = gsub("MHPA", "HMPAA", Compound)) %>% 
       mutate(Compound = gsub("MAPH", "MAP",  Compound))
```

```{r, warning=F, echo=F, message=F}
metadata <- left_join(metadata, database, by = c("Strain" = "Strain_ID"))
```

Quantification & calculations: using a standard curve the samples are quantified

```{r, warning=F, echo=F, message=F}
met_Std_11 <- met %>% 
  filter(grepl('20211007_MH_LT_Bact_PMStds11x', Name)) %>% 
  # mutate(Type = as.factor(Type)) %>% 
  filter(Type %in% "Standard") %>% 
  filter(Std_conc != "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
met_Zero <- met %>% 
  filter(grepl('20211007_MH_LT_Bact_PMStds11x', Name)) %>% 
  # filter(Type %in% "Zero") %>% 
  filter(Std_conc %in% "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
met_Std_11 <- rbind(met_Std_11, met_Zero)
met_Std_11 %<>% filter(!Std_conc %in% "NA")
```

```{r, warning=F, echo=F, message=F}
met_Std_11 %>% 
  filter(Compound != "HMPAA") %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x= Std_conc, y = Area)) + 
  geom_point(aes(colour = Compound), show.legend = FALSE) +
  geom_smooth(method = "lm", color="black", formula = y ~ 0 + x) +
  stat_poly_eq(formula = y ~ 0 + x,
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                parse = TRUE, label.y.npc = "top", label.x.npc = "left") +
  stat_fit_glance(method = "lm", method.args = list(formula = y ~ 0 + x),
                      geom = 'text', aes(label = paste("P-val. = ", 
                      signif(..p.value.., digits = 4), sep = "")), label.y = "bottom", label.x = "center") +
  facet_wrap(~Compound, scales = "free") +
  labs(y = "Area",
       x = "Concentration")
```

```{r, warning=F, echo=F, message=F}
St_slopes_11 <- met_Std_11 %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>%
  group_by(Compound) %>% 
  do(lm(Area ~ 0 + Std_conc,  data = .) %>%  coef() %>%   as.data.frame() %>% slice(1))
```

```{r, warning=F, echo=F, message=F}
colnames(St_slopes_11) <- c("Compound", "Slope")
```

```{r, warning=F, echo=F, message=F}
met_Std_HMPAA <- met %>% 
  filter(grepl('20211007_MH_LT_Bact_MHPA', Name)) %>% 
  # mutate(Type = as.factor(Type)) %>% 
  filter(Type %in% "Standard") %>% 
  filter(Std_conc != "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
met_Zero_HMPAA <- met %>% 
  filter(grepl('20211007_MH_LT_Bact_MHPA', Name)) %>% 
  # filter(Type %in% "Zero") %>% 
  filter(Std_conc %in% "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
met_Std_HMPAA <- rbind(met_Std_HMPAA, met_Zero_HMPAA)
met_Std_HMPAA %<>% filter(!Std_conc %in% "NA")
```

```{r, warning=F, echo=F, message=F}
met_Std_HMPAA %>% 
  filter(Compound %in% "HMPAA") %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x= Std_conc, y = Area)) + 
  geom_point(aes(colour = Compound), show.legend = FALSE) +
  geom_smooth(method = "lm", color="black", formula = y ~ 0 + x) +
  stat_poly_eq(formula = y ~ 0 + x,
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                parse = TRUE, label.y.npc = "top", label.x.npc = "left") +
  stat_fit_glance(method = "lm", method.args = list(formula = y ~ 0 + x),
                      geom = 'text', aes(label = paste("P-val. = ", 
                      signif(..p.value.., digits = 4), sep = "")), label.y = "bottom", label.x = "center") +
  facet_wrap(~Compound, scales = "free") +
  labs(y = "Area",
       x = "Concentration")
```

```{r, warning=F, echo=F, message=F}
St_slopes_HMPAA <- met_Std_HMPAA %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>%
  group_by(Compound) %>% 
  do(lm(Area ~ 0 + Std_conc,  data = .) %>%  coef() %>%   as.data.frame() %>% slice(1))
```

```{r, warning=F, echo=F, message=F}
colnames(St_slopes_HMPAA) <- c("Compound", "Slope")
```

```{r, warning=F, echo=F, message=F}
St_slopes_11_HMPAA <- left_join(St_slopes_11, St_slopes_HMPAA, by = "Compound")

St_slopes_11_HMPAA_unite <- St_slopes_11_HMPAA %>% unite("slope_all", Slope.x:Slope.y, remove = FALSE)
St_slopes_11_HMPAA_unite$slope_all <- gsub("_NA", "", St_slopes_11_HMPAA_unite$slope_all)
St_slopes_11_HMPAA_unite$slope_all <- gsub("NA_", "", St_slopes_11_HMPAA_unite$slope_all)
St_slopes_11_HMPAA_unite %<>% dplyr::select(Compound, slope_all)
```

```{r, warning=F, echo=F, message=F}
met %<>% filter(Compound != "AAPO") 
```

```{r, warning=F, echo=F, message=F}
met$Name <- gsub("20211007_MH_LT_Bact_", "", met$Name)
met_t0  <- met %>% filter(Name %in% c("t0_MBOA", "t0_DMSO"))
met %<>% filter(Type %in% "Analyte")
met <- rbind(met, met_t0)
met <- left_join(met, mol, by = "Compound")
met %<>% mutate(M = as.numeric(M)) # NAs can not be mutated to numeric, a small error message pops up
```

```{r, warning=F, echo=F, message=F}
metadata %<>% dplyr::rename(Name = Sample_ID)
met <- left_join(met, metadata, by = "Name")
# met %<>% mutate(Dilution = as.numeric(Dilution))
met %<>%
  mutate(Strain = case_when(Name == "t0_MBOA" ~ "t0",
                            Name == "t0_DMSO" ~ "t0",
                            Strain != "t0_DMSO" & Strain != "t0_MBOA" ~ Strain)) %>% 
  mutate(Treatment = case_when(Name == "t0_MBOA" ~ "MBOA",
                               Name == "t0_DMSO" ~ "DMSO",
                               Treatment != "t0_DMSO" & Treatment != "t0_MBOA" ~ Treatment))

met %<>% as.data.frame()
```

```{r, warning=F, echo=F, message=F}
met$Treatment <- factor(met$Treatment, levels = c("DMSO", "MBOA"))
```

```{r, warning=F, echo=F, message=F}
met <- left_join(met, St_slopes_11_HMPAA_unite, by = "Compound")
```

```{r, warning=F, echo=F, message=F}
met$cols <- as.character(met$Compound)
met[met$Compound == "MBOA-Glc", ]$cols <- "cornflowerblue"
met[met$Compound == "HMPMA", ]$cols <-  "darkolivegreen4"


met[met$Compound == "AMPO", ]$cols <- "darkred"
met[met$Compound == "AAMPO", ]$cols <- "firebrick1"
met[met$Compound == "HMPAA", ]$cols <- "gold2"
met[met$Compound == "MBOA", ]$cols <- "darkblue"

met[met$Compound == "HMBOA", ]$cols <- "skyblue"
met[met$Compound == "BOA", ]$cols <- "slateblue"


met[met$Compound == "DIMBOA", ]$cols <-  "tan"
met[met$Compound == "APO", ]$cols <-  "pink"
# met[met$Compound == "AAPO", ]$cols <-  "palevioletred"

temp <- data.frame(met$Compound, met$cols)
temp <- plyr::ddply(temp, .variables="met.Compound", .fun=unique)   #library(plyr)
level_cols_Compound <- as.character(temp[,2])
names(level_cols_Compound) <- temp[,1]
```

```{r, warning=F, echo=F, message=F}
# met$Area_sample <- as.numeric(met$Area) / 0.025 

met$Area_sample <- as.numeric(met$Area) / 0.04 
met$slope_all <- as.numeric(met$slope_all)
met$ng_sample <- met$Area_sample / met$slope_all 
met$uMol <- (met$ng_sample) / met$M
```

Explore raw data

```{r, warning=F, echo=F, message=F}
met %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  ggplot(aes(x = Strain, y = uMol, group = Treatment)) +
  geom_point(aes(colour = Treatment), show.legend = TRUE) +
  facet_wrap(~ Compound, scales = "free") +
  scale_colour_manual(values = c("darkblue", "darkred"))+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) 
```

Stacked Bargraph

```{r, warning=F, echo=F, message=F}
met %>% 
  # mutate(Compound = as.factor(Compound)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  filter(Treatment %in% "MBOA") %>% 
  # filter(Strain %in% c("SCS", "SCR")) %>% 
  # filter(Treatment %in% "MBOA_high") %>% 
  ggplot(aes(x = Strain, y = uMol)) +
  geom_bar(aes(fill = Compound), stat = "identity", size = 2, show.legend = TRUE) +
  # geom_path(aes(colour = Compound), stat = "summary") +
  # facet_wrap(~ Strain, scales = "fixed") +
  scale_fill_manual(values = level_cols_Compound)+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) +
  labs(y = "concentration [µM]" )
```

## Phylogenetic tree

```{r, warning=F, echo=F, message=F}
### Newick tree ###
species_tree <- ape::read.tree("Input/SpeciesTree_rooted.txt")

metadata_tree <- metadata %>% mutate(Host_plant = Collection) %>% 
  mutate(Host_plant = gsub("Species_collection", "maize", Host_plant)) %>% 
  mutate(Host_plant = gsub("Maize_collection", "maize", Host_plant)) %>% 
  mutate(Host_plant = gsub("MBOA", "maize", Host_plant)) %>% 
  mutate(Host_plant = gsub("Compartment_collection", "maize", Host_plant)) %>% 
  mutate(Host_plant = gsub("AtSphere", "arabidopsis", Host_plant)) %>% 
  mutate(Host_plant = case_when(Strain == "LBN7" ~ "brassica", Strain != "LBN7" ~ Host_plant)) %>% 
  mutate(Host_plant = case_when(Strain == "LTA6" ~ "wheat", Strain != "LTA6" ~ Host_plant)) %>% 
  mutate(Host_plant = case_when(Strain == "LMS4" ~ "medicago", Strain != "LMS4" ~ Host_plant)) %>% 
  # mutate(Host_plant = case_when(Strain == "KHBO19" ~ "clover", Strain != "KHB019" ~ Host_plant)) %>% 
  dplyr::select(Strain, AMPO_on_MBOA_plates, Host_plant) %>% 
  filter(!Strain %in% NA) %>% 
  filter(!Strain %in% "NA") %>% 
  mutate(AMPO_on_MBOA_plates = gsub("no", "negative", AMPO_on_MBOA_plates)) %>% 
  mutate(AMPO_on_MBOA_plates = gsub("yes", "positive", AMPO_on_MBOA_plates)) %>% 
  add_row(Strain = "KHB019", AMPO_on_MBOA_plates = "negative", Host_plant = "clover") %>%  mutate(Host_plant = factor(Host_plant, levels = c("maize", "arabidopsis", "brassica", "medicago", "wheat", "clover", "NA")))
  
metadata_tree_2 <- metadata_tree %>% 
        mutate(Host_plant = gsub("brassica", "other", Host_plant)) %>% 
        mutate(Host_plant = gsub("medicago", "other", Host_plant)) %>% 
        mutate(Host_plant = gsub("wheat", "other", Host_plant)) %>% 
        mutate(Host_plant = gsub("clover", "other", Host_plant))

p_host_simple <-  species_tree %>% 
      ggtree::ggtree(branch.length = "branch.length", ladderize = TRUE) %<+% metadata_tree_2 +
      geom_tippoint(aes(color=Host_plant), size = 3) + 
      geom_tiplab(align=TRUE, linetype='dashed', linesize=.3, size = 3, offset = 0.02) +
      scale_color_manual(values = c("#3e5437", "tan3", "grey"), labels = c("maize", "arabidopsis", "other"), name = "Host plant") +
      xlim(0, 2) 

p_host_simple
```

```{r, warning=F, echo=F, message=F}
# add orthogroup data

met_OG <- met %>% mutate(OG0001785 = 0,
               OG0002141 = 0,
               OG0002970 = 0,
               OG0002971 = 0,
               OG0002972 = 0)

met_OG %<>%
  mutate(OG0001785 = case_when(AMPO_on_MBOA_plates == "yes" ~ "2",
                            AMPO_on_MBOA_plates == "no" ~ "0")) %>% 
  mutate(OG0001785 = case_when(Strain == "LWH7" ~ "4",
                            Strain != "LWH7" ~ OG0001785)) %>%
  mutate(OG0001785 = case_when(Strain == "LWH3" ~ "3",
                            Strain != "LWH3" ~ OG0001785)) %>%
  mutate(OG0001785 = case_when(Strain == "LWS13" ~ "3",
                            Strain != "LWS13" ~ OG0001785)) %>%
  mutate(OG0002141 = case_when(AMPO_on_MBOA_plates == "yes" ~ "2",
                            AMPO_on_MBOA_plates == "no" ~ "0")) %>% 
  mutate(OG0002141 = case_when(Strain == "LWH7" ~ "3",
                            Strain != "LWH7" ~ OG0002141)) %>%
  mutate(OG0002141 = case_when(Strain == "LWH3" ~ "3",
                            Strain != "LWH3" ~ OG0002141)) %>%
  mutate(OG0002141 = case_when(Strain == "LWS13" ~ "3",
                            Strain != "LWS13" ~ OG0002141)) %>%
  mutate(OG0002141 = case_when(Strain == "LWH13" ~ "1",
                            Strain != "LWH13" ~ OG0002141)) %>%
  mutate(OG0002141 = case_when(Strain == "LM3X" ~ "1",
                            Strain != "LM3X" ~ OG0002141)) %>%
  mutate(OG0002141 = case_when(Strain == "LMB2" ~ "1",
                            Strain != "LMB2" ~ OG0002141)) %>%
  mutate(OG0002141 = case_when(Strain == "LMX7" ~ "1",
                            Strain != "LMX7" ~ OG0002141)) %>%
  mutate(OG0002141 = case_when(Strain == "LWO14" ~ "1",
                            Strain != "LWO14" ~ OG0002141)) %>%
  mutate(OG0002970 = case_when(AMPO_on_MBOA_plates == "yes" ~ "1",
                           AMPO_on_MBOA_plates == "no" ~ "0")) %>% 
  mutate(OG0002971 = case_when(AMPO_on_MBOA_plates == "yes" ~ "1",
                            AMPO_on_MBOA_plates == "no" ~ "0")) %>% 
  mutate(OG0002972 = case_when(AMPO_on_MBOA_plates == "yes" ~ "1",
                            AMPO_on_MBOA_plates == "no" ~ "0"))  
  
```

```{r, warning=F, echo=F, message=F}
tax_host_pheno <- ggtree::gheatmap(p_host_simple + new_scale_fill(),
         metadata_tree %>% 
         dplyr::select(Strain, AMPO_on_MBOA_plates) %>% 
           dplyr::rename('Plate assay' = AMPO_on_MBOA_plates) %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain"),
         offset=0.3,
         width=0.2,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("lightgoldenrod", "darkred")) +
  labs(fill = "AMPO phenotype")

tax_host_pheno
```

```{r, warning=F, echo=F, message=F}
MBOA_AMPO_met_tax_host <-ggtree::gheatmap(tax_host_pheno + new_scale_fill(),
         met %>% filter(Compound %in% c("MBOA", "AMPO")) %>% 
           filter(Treatment %in% "MBOA") %>% 
           filter(!uMol %in% 0.00000000) %>%
           filter(!uMol > 600) %>%
           mutate(Compound = factor(Compound, levels = c("MBOA", "AMPO", "AAMPO", "HMPAA"))) %>% 
         # filter(MBOA_DMSO_per < 0) %>% 
         # mutate(Chem = as.character(Conc)) %>%  
         dplyr::select(Strain, Compound, uMol) %>% tidyr::spread(Compound, uMol) %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain") ,
         offset=0.5,
         width=0.4,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_gradient2(low = "grey50", midpoint = 20, high = "darkred")+
  labs(#x = "Benzoxazinoid",
       #y = "Relative AUC", 
       colour = "AMPO phenotype",
       shape = "Host plant",
       fill = "Concentration [uM]") 

MBOA_AMPO_met_tax_host
#dev.off()
```

```{r, warning=F, echo=F, message=F}
met_OG_selected <- met_OG %>% 
  # filter(Strain != "t0") %>% 
  #          mutate(Strain = gsub("NBC", "KHB019",  Strain)) %>% 
         dplyr::select(Strain, OG0001785, OG0002141, OG0002970, OG0002971, OG0002972) %>% 
           mutate(OG0001785 = replace_na(OG0001785, "0")) %>% 
           mutate(OG0002141 = replace_na(OG0002141, "0")) %>% 
           mutate(OG0002970 = replace_na(OG0002970, "0")) %>% 
           mutate(OG0002971 = replace_na(OG0002971, "0")) %>% 
           mutate(OG0002972 = replace_na(OG0002972, "0")) %>% 
           add_row(Strain = "KHB019", OG0001785 = "0", OG0002141 = "0", OG0002970 = "0", OG0002971 = "0", OG0002972 = "0") %>% unique()
```


```{r, warning=F, echo=F, message=F}
Fig.4.MBOA_AMPO_met_tax_shape_OG_host <- ggtree::gheatmap(MBOA_AMPO_met_tax_host + new_scale_fill(),
         met_OG_selected %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain"),
         offset=0.8,
         width=1,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("gray100", "gray80", "gray60", "gray40", "gray20", "white")) +
  labs(fill = "Orthogroup copies")

Fig.4.MBOA_AMPO_met_tax_shape_OG_host

# ggsave(plot = Fig.4.MBOA_AMPO_met_tax_shape_OG_host, filename = "Fig.4.MBOA_AMPO_met_tax_shape_OG_host.pdf", width = 23, height = 23, dpi = 300, scale = 1, units = "cm")
# 
# ggsave(plot = Fig.4.MBOA_AMPO_met_tax_shape_OG_host, filename = "Fig.4.MBOA_AMPO_met_tax_shape_OG_host.svg", width = 23, height = 23, dpi = 300, scale = 1, units = "cm")
```

```{r, warning=F, echo=F, message=F}
met_bar_phylo <- met %>% filter(Treatment %in% "MBOA") %>% filter(!Compound %in% c("AMPO", "AAMPO")) %>% dplyr::select(Strain, Compound, uMol) %>% pivot_wider(names_from = Compound, values_from = uMol) 

met_bar_phylo$Strain <- gsub("NBC", "KHB019", met_bar_phylo$Strain)
met_bar_phylo %<>% as.data.frame()

# met_bar_phylo %<>% column_to_rownames("Strain")

met_phylo <- met %>% filter(Treatment %in% "MBOA") %>% filter(Strain != "t0") %>% dplyr::select(Strain, Compound, uMol)
met_phylo$Strain <- gsub("NBC", "KHB019", met_phylo$Strain)
```

```{r, warning=F, echo=F, message=F}
p_tree <-  species_tree %>% 
      ggtree::ggtree(branch.length = "none", ladderize = TRUE, right = FALSE) %<+% metadata_tree + 
      geom_tippoint(aes(color=AMPO_on_MBOA_plates), show.legend = FALSE) + 
      geom_tiplab(size=3) + 
      xlim(0, 30)+
      scale_color_manual(values = c("gold", "darkred"))
```

```{r, warning=F, echo=F, message=F}
p_met <- met_phylo %>% 
         mutate(uMol = case_when(Strain == "KHB019" ~ 0,
                                  Strain != "KHB019" ~ uMol)) %>% 
        ggplot(aes(y = Strain, x = uMol)) +
        geom_bar(aes(fill = Compound), stat = "identity") +
        # scale_fill_jco()+
        scale_fill_manual(values = level_cols_Compound)+
        theme_tree2() +
        labs(x = "Concentration [uM]")

plot_tax_met_micro <- p_met %>% 
  insert_left(p_host_simple, width = 0.5)  

print(plot_tax_met_micro)

#ggsave(plot = plot_tax_met_micro, filename = "plot_tax_met_micro.pdf", width = 15, height = 15, dpi = 300, scale = 1, units = "cm")
# 
#ggsave(plot = plot_tax_met_micro, filename = "plot_tax_met_micro.svg", width = 15, height = 15, dpi = 300, scale = 1, units = "cm")
```

# DIMBOA-Glc metabolisation
```{r, warning=F, echo=F, message=F}
# Metadata
# metadataDA <- read_excel("220105_Microbacteria_DMSO_DMG_AMPO_Samplelist_metabolites_qTOF.xlsx", sheet = 1) %>% filter(!Treatment %in% "A")
# 
# write.csv(metadataDA, "Microbacteria_DMSO_DMG_sample_list.csv")

metadataDA <- read.csv("Input/Microbacteria_DMSO_DMG_sample_list.csv")

metadataDA <- left_join(metadataDA, database, by = c("Strain" = "Strain_ID"))
# Metabolite data
# metDA <- read_excel("20220203_MH_LT_Microbact_DMG_metabolites_quantified_xlsx.xlsx", sheet = 2)
#   
#   metDA %<>% mutate(Treat = case_when(grepl("_A", Name) ~ "AMPO",
#                                   grepl("_D", Name) ~ "DMSO",
#                                   grepl("_G", Name) ~ "DMG")) %>% filter(!Treat %in% "AMPO")

# write.csv(metDA, "Microbact_DMG_metabolites_quantified.csv")

metDA <- read.csv("Input/Microbact_DMG_metabolites_quantified.csv")


metDA %<>% dplyr::rename(Std_conc = Std..Conc, ng_mL_ML = ng.mL) %>% 
       mutate(Compound = gsub("MHPA", "HMPAA", Compound)) %>% 
       mutate(Compound = gsub("MAPH", "MAP",  Compound))
```


Quantification & calculations using a standard curve the samples are quantified

```{r, warning=F, echo=F, message=F}
metDA_Std_11 <- metDA %>% 
  filter(grepl('20220203_MH_LT_Microbact_metabolites_PM11xStds', Name)) %>% 
  filter(Type %in% "Standard") %>% 
  filter(Std_conc != "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
metDA_Zero <- metDA %>% 
  filter(grepl('20220203_MH_LT_Microbact_metabolites_PM11xStds', Name)) %>% 
  filter(Std_conc %in% "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
metDA_Std_11 <- rbind(metDA_Std_11, metDA_Zero)
metDA_Std_11 %<>% filter(!Std_conc %in% "NA")
```

```{r, warning=F, echo=F, message=F}
metDA_Std_11 %>% 
  filter(!Compound %in% c("AAPO", "HMPAA", "DIBOA-Glc", "HBOA-Glc", "HM2BOA-Glc", "HMBOA-Glc", "MAPH")) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x= Std_conc, y = Area)) + 
  geom_point(aes(colour = Compound), show.legend = FALSE) +
  geom_smooth(method = "lm", color="black", formula = y ~ 0 + x) +
  stat_poly_eq(formula = y ~ 0 + x,
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                parse = TRUE, label.y.npc = "top", label.x.npc = "left") +
  stat_fit_glance(method = "lm", method.args = list(formula = y ~ 0 + x),
                      geom = 'text', aes(label = paste("P-val. = ", 
                      signif(..p.value.., digits = 4), sep = "")), label.y = "bottom", label.x = "center") +
  facet_wrap(~Compound, scales = "free") +
  labs(y = "Area",
       x = "Concentration")
```

```{r, warning=F, echo=F, message=F}
St_slopes_11_DA <- metDA_Std_11 %>% 
  filter(!Compound %in% c("AAPO", "HMPAA", "DIBOA-Glc", "HBOA-Glc", "HM2BOA-Glc", "HMBOA-Glc", "MAP")) %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>%
  group_by(Compound) %>% 
  do(lm(Area ~ 0 + Std_conc,  data = .) %>%  coef() %>%   as.data.frame() %>% slice(1))
```

```{r, warning=F, echo=F, message=F}
colnames(St_slopes_11_DA) <- c("Compound", "Slope")
```

```{r, warning=F, echo=F, message=F}
metDA_Std_HMPAA <- metDA %>% 
  filter(grepl('20220203_MH_LT_Microbact_metabolites_MHPA', Name)) %>% 
  # mutate(Type = as.factor(Type)) %>% 
  filter(Type %in% "Standard") %>% 
  filter(Std_conc != "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
metDA_Zero_HMPAA <- metDA %>% 
  filter(grepl('20220203_MH_LT_Microbact_metabolites_MHPA', Name)) %>% 
  # filter(Type %in% "Zero") %>% 
  filter(Std_conc %in% "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
metDA_Std_HMPAA <- rbind(metDA_Std_HMPAA, metDA_Zero_HMPAA)
metDA_Std_HMPAA %<>% filter(!Std_conc %in% "NA")
```

```{r, warning=F, echo=F, message=F}
metDA_Std_HMPAA %>% 
  filter(Compound %in% "HMPAA") %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x= Std_conc, y = Area)) + 
  geom_point(aes(colour = Compound), show.legend = FALSE) +
  geom_smooth(method = "lm", color="black", formula = y ~ 0 + x) +
  stat_poly_eq(formula = y ~ 0 + x,
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                parse = TRUE, label.y.npc = "top", label.x.npc = "left") +
  stat_fit_glance(method = "lm", method.args = list(formula = y ~ 0 + x),
                      geom = 'text', aes(label = paste("P-val. = ", 
                      signif(..p.value.., digits = 4), sep = "")), label.y = "bottom", label.x = "center") +
  facet_wrap(~Compound, scales = "free") +
  labs(y = "Area",
       x = "Concentration")
```

```{r, warning=F, echo=F, message=F}
St_slopes_HMPAA_DA <- metDA_Std_HMPAA %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>%
  group_by(Compound) %>% 
  do(lm(Area ~ 0 + Std_conc,  data = .) %>%  coef() %>%   as.data.frame() %>% slice(1))
```

```{r, warning=F, echo=F, message=F}
colnames(St_slopes_HMPAA_DA) <- c("Compound", "Slope")
```
join 11 stds and HMPAA
```{r, warning=F, echo=F, message=F}
St_slopes_11_HMPAA_DA_unite <- rbind(St_slopes_11_DA, St_slopes_HMPAA_DA %>% filter(Compound %in% "HMPAA"))

# St_slopes_HMPAA_DA_unite <- St_slopes_HMPAA_DA %>% unite("slope_all", Slope.x:Slope.y, remove = FALSE)
# St_slopes_11_HMPAA_DA_unite$slope_all <- gsub("_NA", "", St_slopes_11_HMPAA_DA_unite$slope_all)
# St_slopes_11_HMPAA_DA_unite$slope_all <- gsub("NA_", "", St_slopes_11_HMPAA_DA_unite$slope_all)
# St_slopes_11_HMPAA_DA_unite %<>% dplyr::select(Compound, slope_all)
```

remove the compounds which are not reliable quantified
```{r, warning=F, echo=F, message=F}
metDA %<>% filter(Compound != "AAPO") 
```

```{r, warning=F, echo=F, message=F}
metDA$Name <- gsub("20220203_MH_LT_Microbact_metabolites_", "", metDA$Name)
# metDA_t0  <- metDA %>% filter(Name %in% c("t0_MBOA", "t0_DMSO"))
metDA %<>% filter(Type %in% "Analyte")
#metDA <- rbind(metDA, metDA_t0)
metDA <- left_join(metDA, mol, by = "Compound")
metDA %<>% mutate(M = as.numeric(M)) # NAs can not be mutated to numeric, a small error message pops up
```

```{r, warning=F, echo=F, message=F}
metadataDA %<>% dplyr::rename(Name = Sample)

metDA <- left_join(metDA, metadataDA, by = "Name")
# met %<>% mutate(Dilution = as.numeric(Dilution))
# metDA %<>%
#   mutate(Strain = case_when(Name == "t0_MBOA" ~ "t0",
#                             Name == "t0_DMSO" ~ "t0",
#                             Strain != "t0_DMSO" & Strain != "t0_MBOA" ~ Strain)) %>% 
#   mutate(Treatment = case_when(Name == "t0_MBOA" ~ "MBOA",
#                                Name == "t0_DMSO" ~ "DMSO",
#                                Treatment != "t0_DMSO" & Treatment != "t0_MBOA" ~ Treatment))
# 
# metDA %<>% as.data.frame()
```

```{r, warning=F, echo=F, message=F}
# met$Strain <- factor(met$Strain, levels = c("LSP13",  "LMD1", "LME3", "LMB2", "LMI1x",
#                                             "LPD2", "LMX9", "LAC11", "LBA112", "LRC7.O", "NBC",
#                                             "LMG1", "LMN1", "LST17", "LBA21", "SCR", "SCS"))

metDA$Treatment <- factor(metDA$Treatment, levels = c("D", "G", "A"))
```

```{r, warning=F, echo=F, message=F}
metDA <- left_join(metDA, St_slopes_11_HMPAA_DA_unite, by = "Compound")
```

```{r, warning=F, echo=F, message=F}
metDA$cols <- as.character(metDA$Compound)
metDA[metDA$Compound == "AMPO", ]$cols <- "darkred"
metDA[metDA$Compound == "AAMPO", ]$cols <- "firebrick1"
metDA[metDA$Compound == "HMPAA", ]$cols <- "gold2"
metDA[metDA$Compound == "MBOA", ]$cols <- "darkblue"
metDA[metDA$Compound == "MAP", ]$cols <- "khaki"

metDA[metDA$Compound == "HMBOA", ]$cols <- "skyblue"
metDA[metDA$Compound == "BOA", ]$cols <- "slateblue"
metDA[metDA$Compound == "MBOA-Glc", ]$cols <- "darkmagenta"


metDA[metDA$Compound == "HMPMA", ]$cols <-  "slategray"
metDA[metDA$Compound == "DIMBOA", ]$cols <-  "tan"
metDA[metDA$Compound == "APO", ]$cols <-  "pink"
# metDA[metDA$Compound == "AAPO", ]$cols <-  "pink3"

metDA[metDA$Compound == "DIMBOA-Glc", ]$cols <-  "darkorange2"
metDA[metDA$Compound == "HDMBOA-Glc", ]$cols <-  "thistle3"
metDA[metDA$Compound == "HBOA-Glc", ]$cols <-  "slategray"
metDA[metDA$Compound == "DIBOA-Glc", ]$cols <-  "slategray"
metDA[metDA$Compound == "HMBOA-Glc", ]$cols <-  "skyblue2"
metDA[metDA$Compound == "HM2BOA-Glc", ]$cols <-  "skyblue4"
metDA[metDA$Compound == "DIM2BOA-Glc", ]$cols <-  "sandybrown"
metDA[metDA$Compound == "HDM2BOA-Glc", ]$cols <-  "thistle1"

temp <- data.frame(metDA$Compound, metDA$cols)
temp <- plyr::ddply(temp, .variables="metDA.Compound", .fun=unique)   #library(plyr)
level_cols_Compound <- as.character(temp[,2])
names(level_cols_Compound) <- temp[,1]
```


Dilutions: 
150 + 350 = 500 
50 + 700 = 750 
500 / 150 = 3.33333
750 / 50 = 15

1/((500/150)*15) = 0.02

for some reason the MBOA concentrations are a bit too high (factor 1.273141 controlled with t0 MBOA sample) --> ev because of precipitaion issue?

```{r, warning=F, echo=F, message=F}
metDA$Area_sample <- as.numeric(metDA$Area) / 0.025 
# metDA$Slope <- as.numeric(metDA$Slope)
metDA$ng_sample <- metDA$Area_sample / metDA$Slope 
metDA$uMol <- (metDA$ng_sample) / metDA$M
```

Explore raw data
```{r, warning=F, echo=F, message=F}
metDA %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  ggplot(aes(x = Strain, y = uMol, group = Treatment)) +
  geom_point(aes(colour = Treatment), show.legend = TRUE) +
  facet_wrap(~ Compound, scales = "free") +
  scale_colour_manual(values = c("darkblue", "darkorange2", "maroon"))+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) 
```

```{r, warning=F, echo=F, message=F}
metDA %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x = Strain, y = uMol, group = Treatment)) +
  geom_bar(aes(fill = Treatment), stat = "identity", size = 2, show.legend = TRUE) +
  facet_wrap(~ Compound, scales = "fixed") +
  scale_fill_manual(values = c("darkblue", "darkorange2", "maroon"))+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) 
```

```{r, warning=F, echo=F, message=F}
metDA %>% 
  filter(!Compound %in% c("DIM2BOA-Glc", "HDM2BOA-Glc")) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  filter(Treatment %in% "G") %>% 
  ggplot(aes(x = Strain, y = uMol)) +
  geom_bar(aes(fill = Compound), stat = "identity", size = 2, show.legend = TRUE) +
  scale_fill_manual(values = level_cols_Compound)+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) +
  labs(title = "DIMBOA-Glc")
```

## Phylogenetic tree
```{r, warning=F, echo=F, message=F}
### Newick tree ###
species_tree <- ape::read.tree("Input/SpeciesTree_rooted.txt")

metadata_tree <- metadata %>% dplyr::select(Strain, AMPO_on_MBOA_plates) %>% filter(!Strain %in% NA)

p <-  species_tree %>% 
      ggtree::ggtree(branch.length = "none", ladderize = TRUE) %<+% metadata_tree + 
      geom_tippoint(aes(color=AMPO_on_MBOA_plates)) + 
      geom_tiplab(size=3) + 
      scale_color_manual(values = c("gold", "darkred"))
```

```{r, message=F, echo=F, warning=F}
ggtree::gheatmap(p,
         metDA %>% filter(Compound %in% c("DIMBOA-Glc")) %>% 
           filter(Treatment %in% "G") %>% 
         dplyr::select(Strain, uMol) %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain") ,
         offset=0.1,
         width=0.2,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_gradient2(low = "white", mid = "gold", high ="darkred", midpoint = 500) +
  labs(title = "MBOA concentration",
       subtitle = "all strains") +
 guides(color=FALSE) #+
```

```{r, warning=F, echo=F, message=F}
metDA_bar_phylo <- metDA %>% filter(Treatment %in% "G") %>% dplyr::select(Strain, Compound, uMol) %>% pivot_wider(names_from = Compound, values_from = uMol) 

metDA_bar_phylo$Strain <- gsub("NBC", "KHB019", metDA_bar_phylo$Strain)
metDA_bar_phylo %<>% as.data.frame()

metDA_phylo <- metDA %>% filter(Treatment %in% "G") %>% filter(Strain != "T0") %>% filter(!Compound %in% c("HDM2BOA-Glc", "DIM2BOA-Glc")) %>% dplyr::select(Strain, Compound, uMol)
metDA_phylo$Strain <- gsub("NBC", "KHB019", metDA_phylo$Strain)
```

```{r, warning=F, echo=F, message=F}
p_metDA_DMG <- metDA_phylo %>% 
        ggplot(aes(y = Strain, x = uMol)) +
        geom_bar(aes(fill = Compound), stat = "identity") +
        xlim(0, 800)+
        scale_fill_manual(values = level_cols_Compound)+
        theme_tree2() +
        labs(x = "Concentration [uM]")

plot_tax_metDA_DMG_micro <- p_metDA_DMG %>% 
  insert_left(p_host_simple, width = 0.5)  


plot_tax_metDA_DMG_micro

# ggsave(plot = plot_tax_metDA_DMG_micro, filename = "plot_tax_metDA_DMG_micro_NBC.pdf", width = 15, height = 15, dpi = 300, scale = 1, units = "cm")
#  
# ggsave(plot = plot_tax_metDA_DMG_micro, filename = "plot_tax_metDA_DMG_micro_NBC.svg",width = 15, height = 15, dpi = 300, scale = 1, units = "cm")
```


# Metabolite categories for comparative genomics

```{r, warning=F, echo=F, message=F}
met_all <- rbind(met %>% dplyr::select(-Nr, -Nr.x, -T, -n, -Nr.y) %>% dplyr::rename(Slope = slope_all, IS.Area = 'IS Area', Detection.Flags = 'Detection Flags'), metDA %>% dplyr::select(-X.1, -X.x, -X., -random, -Nr.x, -Nr.y, -X0, -Treat, -X.y) %>% dplyr::rename('%Dev' = X.Dev))
```

```{r, warning=F, echo=F, message=F}
met_all_quant <- met_all %>% dplyr::select(Strain, Treatment, Compound, uMol) %>% filter(Treatment != "DMSO") %>% mutate(Treatment_Compound = interaction(Treatment, Compound, sep = "_")) %>% dplyr::select(-Treatment, -Compound) %>% filter(!Strain %in% c("t0", "T0")) %>% filter(Treatment_Compound != "MBOA_AAMPO") %>% unique() %>% replace(is.na(.), 0) %>%  pivot_wider(names_from = Treatment_Compound, values_from = uMol, values_fn = mean)
```

**MBOA**
```{r, warning=F, echo=F, message=F}
## MBOA metabolization < 400
max_MBOA <- met_all_quant$MBOA_MBOA %>% max()
# [1] 825.9893

## AMPO formation > 2.5
max_AMPO <-met_all_quant$MBOA_AMPO %>% max()
# [1] 825.9893

## HMPAA formation > 25
max_HMPAA <-met_all_quant$MBOA_HMPAA %>% max()
# [1] 70.50679

tibble(max_MBOA, max_AMPO, max_HMPAA) %>% knitr::kable(caption = "MBOA metabolisation")
```

**DIMBOA-Glc metabolization**
```{r, warning=F, echo=F, message=F}
## DIMBOA-Glc metabolization < 400 (LMI1x 322 uM)
max_DIMBOA_Glc <- met_all_quant$`G_DIMBOA-Glc` %>% max()
# [1] 575.2392

## MBOA formation > 40 (LMI1x 43.7 uM)
max_MBOA <- met_all_quant$G_MBOA %>% max()
# [1] 120.9501

## AMPO formation DMG > 2.5
max_AMPO <- met_all_quant$G_AMPO %>% max()
# [1] 7.355668

tibble(max_DIMBOA_Glc, max_MBOA, max_AMPO) %>% knitr::kable(caption = "DIMBOA-Glc metabolisation")
```

**Groups**
```{r, warning=F, echo=F, message=F}
met_all_quali <- met_all_quant %>%
  mutate(MBOA_met = case_when(MBOA_MBOA > 400 ~ FALSE,
                              MBOA_MBOA < 400 ~ TRUE)) %>%
  mutate(MBOA_AMPO = case_when(MBOA_AMPO > 2.5 ~ TRUE,
                               MBOA_AMPO < 2.5 ~ FALSE)) %>%
  mutate(MBOA_HMPAA = case_when(MBOA_HMPAA > 15 ~ TRUE,
                              MBOA_HMPAA < 15 ~ FALSE)) %>%
  mutate(DIMBOA_Glc_met = case_when(`G_DIMBOA-Glc` > 400 ~ FALSE,
                              `G_DIMBOA-Glc` < 400 ~ TRUE)) %>%
  mutate(DMG_MBOA = case_when(G_MBOA > 40 ~ TRUE,
                              G_MBOA < 40 ~ FALSE)) %>%
  mutate(DMG_AMPO = case_when(G_AMPO > 2.5 ~ TRUE,
                              G_AMPO < 2.5 ~ FALSE)) %>%
  dplyr::select(Strain, MBOA_met, MBOA_AMPO, MBOA_HMPAA, DIMBOA_Glc_met, DMG_MBOA, DMG_AMPO) %>%    left_join(., database %>% dplyr::select(Strain_ID, AMPO_on_MBOA_plates), by = c("Strain" = "Strain_ID"))
```

# GROWTH MICROBACTERIA: TSB (Metabolite samples) & MM (Carbon source)
```{r, include=FALSE}
Strain_Data <- read_excel("Input/210513_experimental_design_Microbacteria_MRB_At_metabolites.xlsx", sheet = 5)
Strain_Data 
database <- read_excel("Input/210513_experimental_design_Microbacteria_MRB_At_metabolites.xlsx", sheet = 2)
Strain_Data <- left_join(Strain_Data, database, by = c("Strain" = "Strain_ID")) %>% dplyr::select(-'old name', -'Possible_species/best_hit') 
Strain_Data$Well_Plate <- interaction(Strain_Data$Well, Strain_Data$Plate, sep = "_")

Strain_Data %<>% arrange(., desc(Strain))
Strain_Data$Block <- rep(c(1:3), times = 64)
```

```{r, warning=FALSE, include=FALSE}
results1 <- data_frame()

# Provide a lookup table for plate information, treatments are spread accross sheets in the table to be read.
# in the first column describe the plates and in the second column the experiment (rep(times = # plates - 2))
treatlist <- data_frame("Treat" =
              c("Dummy",
                "Glc_30000_MM_B_2",
                "Glc_30000_MM_A_2",
                "Glc_30000_MM_B_1",
                "Glc_30000_MM_A_1",
                "MBOA_MM_B_2",
                "MBOA_MM_A_2",
                "MBOA_MM_B_1",
                "MBOA_MM_A_1",
                "DMSO_MM_B_2",
                "DMSO_MM_A_2",
                "DMSO_MM_B_1",
                "DMSO_MM_A_1",
                "MBOA_TSB_B_2",
                "MBOA_TSB_A_2",
                "MBOA_TSB_B_1",
                "MBOA_TSB_A_1",
                "DMSO_TSB_B_2",
                "DMSO_TSB_A_2",
                "DMSO_TSB_B_1",
                "DMSO_TSB_A_1",
                "no_TSB_MM_no_1",
                "Dummy"),
 "Replicate" = c("NULL", rep(c("2", "2","1", "1"), times = 5), "1", "NULL"), # times = all plates but without the dummy plates
 "Plate" = c("NULL", rep(c("B", "A"), times = 10), "NULL", "NULL"), 
# times = all plates but without the dummy plates
 "Media" = c("NULL", rep("MM", times = 12), rep ("TSB", times =8), "TSB_MM", "NULL"),
"Chem" = c("NULL", rep("Glucose", times = 4), rep("MBOA", times = 4), rep("DMSO", times = 4), rep("MBOA", times = 4), rep("DMSO", times = 4), "no", "NULL"), 
"Run" = c("NULL", rep("R1", times = 21), "NULL"))


#Initialize / clear df

#Loop through sheets.

for (i in 1:nrow(treatlist)) {
  tmp <- read_excel("Input/Microbacteria_metabolite_210826_repe_data_reader.xlsx", sheet = i, range = "B26:CU70") %>%   # change for duration runs
    dplyr::select(-`T° 600`) %>% # Drop Temperature
    #pivot_longer(names_to = "Well", values_to = "Density") %>%  #Make long
    gather(key = "Well", value = "Density", 2:ncol(.)) %>%
    mutate(Treat = as.character(treatlist[i,1]), #Add treatment
    Replicate = as.character(treatlist[i,2]),
    Plate = as.character(treatlist[i,3]),
    Media = as.character(treatlist[i,4]),
    Chem = as.character(treatlist[i,5]),
    Run = as.character(treatlist[i,6]),
    Well_Plate = interaction(Well, Plate, sep = "_")) #Add Replicate
results1 <- bind_rows(results1, tmp)
}
```

2nd run was left for 1 cycle longer, therefore the same amount of cycles was kept

```{r, warning=FALSE, include=FALSE}
results2 <- data_frame()

# Provide a lookup table for plate information, treatments are spread accross sheets in the table to be read.
# in the first column describe the plates and in the second column the experiment (rep(times = # plates - 2))
treatlist <- data_frame("Treat" =
              c("Dummy",
                "Glc_30000_MM_B_2",
                "Glc_30000_MM_A_2",
                "Glc_30000_MM_B_1",
                "Glc_30000_MM_A_1",
                "MBOA_MM_B_2",
                "MBOA_MM_A_2",
                "MBOA_MM_B_1",
                "MBOA_MM_A_1",
                "DMSO_MM_B_2",
                "DMSO_MM_A_2",
                "DMSO_MM_B_1",
                "DMSO_MM_A_1",
                "MBOA_TSB_B_2",
                "MBOA_TSB_A_2",
                "MBOA_TSB_B_1",
                "MBOA_TSB_A_1",
                "DMSO_TSB_B_2",
                "DMSO_TSB_A_2",
                "DMSO_TSB_B_1",
                "DMSO_TSB_A_1",
                "no_TSB_MM_no_1",
                "Dummy"),
 "Replicate" = c("NULL", rep(c("2", "2","1", "1"), times = 5), "1", "NULL"), # times = all plates but without the dummy plates
 "Plate" = c("NULL", rep(c("B", "A"), times = 10), "NULL", "NULL"), 
# times = all plates but without the dummy plates
 "Media" = c("NULL", rep("MM", times = 12), rep ("TSB", times =8), "TSB_MM", "NULL"),
"Chem" = c("NULL", rep("Glucose", times = 4), rep("MBOA", times = 4), rep("DMSO", times = 4), rep("MBOA", times = 4), rep("DMSO", times = 4), "no", "NULL"), 
"Run" = c("NULL", rep("R2", times = 21), "NULL"))


#Initialize / clear df

#Loop through sheets.

for (i in 1:nrow(treatlist)) {
  tmp <- read_excel("Input/Microbacteria_metabolite_210930_Run2.xlsx", sheet = i, range = "B26:CU70") %>%   # change for duration runs
    dplyr::select(-`T° 600`) %>% # Drop Temperature
    #pivot_longer(names_to = "Well", values_to = "Density") %>%  #Make long
    gather(key = "Well", value = "Density", 2:ncol(.)) %>%
    mutate(Treat = as.character(treatlist[i,1]), #Add treatment
    Replicate = as.character(treatlist[i,2]),
    Plate = as.character(treatlist[i,3]),
    Media = as.character(treatlist[i,4]),
    Chem = as.character(treatlist[i,5]),
    Run = as.character(treatlist[i,6]),
    Well_Plate = interaction(Well, Plate, sep = "_")) #Add Replicate
results2 <- bind_rows(results2, tmp)
}
```

```{r, warning=FALSE, include=FALSE}
results <- rbind(results1, results2)
```

```{r, warning=FALSE, include=FALSE}
###
###Get times based on difftime. 
results %<>%
  filter(Treat != "Dummy" ) %>%
  filter(Treat != "Dummy") %>%
  filter(Density != "NA") %>%
  group_by(Replicate, Well_Plate, Treat) %>%
  mutate(mins = difftime(Time, Time[1], units = "mins") %>% as.numeric(),
         hrs = difftime(Time, Time[1], units = "hours") %>% as.numeric())  %>% 
  ###Merge with metadata. Well Positions are Part of the stocklist (clean), other metadata comes from the ZmSphere.
  left_join(., Strain_Data, by = "Well_Plate") # 

results %<>% dplyr::select(-Well.y, -Plate.x, -Time, -Row, -Column) %>% dplyr::rename(Well = Well.x, Plate = Plate.y) %>%  as.data.frame() 

results %<>% mutate(Row = Well %>% str_extract(.,"[A-H]"),
                    Column = Well %>% str_extract(.,"[0-9]+") %>% as.numeric()) 

```

```{r, warning=FALSE, include=FALSE}
###Make density increase relative to first timepoint
results %<>%
  filter(!Strain %in% c("Ignore","None")) %>%
  group_by(Replicate, Well_Plate, Treat) %>%
  mutate(Density_increase = Density - Density[1]) %>%
  ungroup %>% as.data.frame()
```

```{r, warning=FALSE, include=FALSE}
results_raw <- results
```

# Growth curves of all plates single treatments
```{r, warning=FALSE, message=F, echo=F}
results %>%  filter(Treat %in% "no_TSB_MM_no_1") %>% 
  ggplot(aes(x=hrs, y= Density_increase)) +
  # stat_summary(geom = "pointrange", fun.y = mean, fun.ymax = max, fun.ymin = min, alpha = 0.3) +
  geom_line(aes(color = Run), size = 1) +
  facet_grid(Row ~ Column, scales = "fixed") +
  theme_bw() +
  # ggtitle("Plate *no bacteria* - check contamination")+
  labs(x = "Time [hours]",
       y = "OD 600", 
       title = "Empty plate",
#       caption = "test",
       subtitle = "control for contamination", 
       color = "Run")
```

```{r, warning=FALSE, message=F, echo=F}
results %>%  
  filter(Media %in% "MM") %>% 
  filter(!Strain %in% "no") %>% 
  filter(!Strain %in% "NBC") %>% 
  ggplot(aes(x=hrs, y= Density_increase)) +
  # stat_summary(geom = "pointrange", fun.y = mean, fun.ymax = max, fun.ymin = min, alpha = 0.3) +
  geom_line(aes(linetype = interaction(Run, Rep, Replicate), color = Chem), size = 1) +
  facet_wrap(~ Strain, scales = "fixed") +
  theme_bw() +
  #scale_color_manual(values = c("cornflowerblue", "cornflowerblue", "darkred", "darkred")) +
  # ggtitle("Plate *no bacteria* - check contamination")+
  labs(x = "Time [hours]",
       y = "OD 600", 
       title = "Growth curves of all strains",
       color = "Run")
```

```{r, warning=FALSE, message=F, echo=F}
results %>%  filter(Media %in% "MM") %>% 
  filter(Chem %in% c("DMSO", "MBOA")) %>% 
  ggplot(aes(x=hrs, y= Density_increase, color = AMPO_on_MBOA_plates)) +
  # stat_summary(geom = "pointrange", fun.y = mean, fun.ymax = max, fun.ymin = min, alpha = 0.3) +
  geom_line(aes(linetype = interaction(Run, Rep, Replicate), color = Chem), size = 1, show.legend = FALSE) +
  facet_wrap( ~ Strain, scales = "fixed") +
  theme_bw() +
  # ggtitle("Plate *no bacteria* - check contamination")+
  labs(x = "Time [hours]",
       y = "OD 600", 
       title = "Variation: Glc_30000_MM_A_1 & Glc_30000_MM_A_2",
#       caption = "test",
       subtitle = "growth in mix - Plate A + Plate B", 
       color = "Run")
```

# Calculate AUC 

To quantify the total bacterial growth over time, we calculate the total area under the curve. This is done with the function "auc()" For comparison between treatments, the total AUC (calculated using density increase) and AUC_raw (calculated using raw density) is normalized with the AUC of the strain grown in the control treatment (no chemicals added, just normal growth media with DMSO). 

```{r, warning=FALSE, include=FALSE, echo=F}
results_AUC <- results %>% 
  filter(Treat != "Dummy") %>% 
  filter(!Strain %in% c("no", NA)) %>% 
  # filter(!Treat %in% c("empty", "glycerol_stock")) %>% 
  group_by(Strain, Well, Chem, Media, Plate, Replicate, Run) %>%
  dplyr::summarize(AUC = auc(hrs, Density_increase), 
            AUC_raw = auc(hrs, Density)) %>%
  ungroup %>% 
  mutate(Chem = as.factor(Chem),
         Plate = as.factor(Plate),
         Replicate = as.factor(Replicate),
         Media = as.factor(Media),
         Well = as.factor(Well),
         Run = as.factor(Run)) %>% 
  mutate(AUC_pos = abs(AUC)) %>% 
  mutate(Strain_Well_Media_Chem_Plate_Replicate_Run = paste(Strain, Well, Media, Chem, Plate, Replicate, Run, sep = "_")) %>%
  #Split by Replicate
  base::split(., .$Strain_Well_Media_Chem_Plate_Replicate_Run) %>% #Make each strain / rep combination its own data.frane
  # mutate(Well_Plate = paste(Well, Plate, sep = "_")) %>% 
  map_dfr(~
  mutate(.x, AUC_norm = AUC_pos / AUC_pos[length(AUC_pos)])) %>%
  #in each df, divide the AUC by the last AUC value; since Control does not start with a number, it is the last treatment
  # dplyr::select(-Replicate) %>% 
  left_join(. , Strain_Data %>% 
  dplyr::select(Strain, Phylum, Family, Genus, AMPO_on_MBOA_plates, Collection) %>% unique(), by = "Strain") %>% 
  # dplyr::select(-Well.y, Plate.y) %>% 
  # dplyr::rename(Well = Well.x, Plate = Plate.x) %>% 
  unique() %>% 
  as.data.frame() #transform into df


## Save table with AUC
#write_rds(results_AUC, paste0("results_AUC.rds"))
```

# Plotting AUC
TSB and Minimal medium

```{r, echo=F, warning=F, message=F, fig.width=20, fig.height=15,}
results_AUC %>%  
  ggplot(aes(x= interaction(as.factor(Chem), Media), y= AUC)) +
  geom_boxplot(aes(fill = AMPO_on_MBOA_plates), position=position_dodge()) + 
  geom_jitter(aes(shape = Run), width = 0.1) + 
  ggpubr::stat_compare_means(label = "p.signif", method = "t.test", ref.group = "DMSO.MM", label.y.npc = 0.8) +
  ylim(0, 60) +
  # stat_summary(fun.data = give.n, geom = "text", fun = median, position = position_fill(0)) +
  #scale_shape_manual(values = shapes_manual ) +
  facet_wrap(~Strain, scales = "fixed") +
  scale_fill_manual(values = c("cornflowerblue", "darkred")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0.5)) +
    labs(y = "AUC", 
         x = "",
         title = "Growth of strains in minimal media",
         colour = "AMPO on plate")
```


### Heatmap growth in minimal media

# TSB: MBOA tolerance

```{r, echo=F, warning=F, message=F, fig.width=20, fig.height=15}
results_AUC %>%  
  filter(Media %in% "TSB") %>% 
  ggplot(aes(x= interaction(as.factor(Chem)), y= AUC)) +
  geom_boxplot(aes(fill = AMPO_on_MBOA_plates), position=position_dodge()) + 
  geom_jitter(width = 0.1) + 
  ggpubr::stat_compare_means(label = "p.signif", method = "t.test", ref.group = "DMSO", label.y = 60) +
  # ylim(0, 80) +
  #scale_shape_manual(values = shapes_manual ) +
  facet_wrap(~Strain, scales = "fixed") +
  scale_fill_manual(values = c("cornflowerblue", "darkred")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0.5)) +
    labs(y = "AUC", 
         title = "Growth of strains in TSB",
         colour = "AMPO on plate")
```

Calculate growth reduction in MM MBOA compared to MM DMSO 
```{r, include=FALSE}
results_AUC_tol_MM <- results_AUC %>% filter(Media %in% "MM") %>% 
  dplyr::select(Strain, Chem, Media, Plate, Replicate, Well, AUC, AMPO_on_MBOA_plates, Run) %>% 
  #mutate(Media_Chem = interaction(Media, Chem)) %>% 
  distinct() %>%  
  pivot_wider(names_from = Chem, values_from = AUC) 

# results_AUC_tol_MM %<>% mutate(Plate_Replicate = (interaction(Plate, Replicate)))

results_AUC_tol_MM$MBOA_DMSO_per <- 100 / (results_AUC_tol_MM$DMSO / results_AUC_tol_MM$MBOA) 

# results_AUC_tol_MM$MBOA_DMSO_diff <- (results_AUC_tol_MM$DMSO / results_AUC_tol_MM$MBOA) * 100
```

```{r, include=FALSE}
results_AUC_tol_MM <- left_join(results_AUC_tol_MM, results_AUC %>% dplyr::select(Strain, Collection), by = "Strain") %>% unique()
```

```{r, echo=F, warning=F, message=F, fig.width=20, fig.height=15}
results_AUC_tol_MM %>%  
  filter(Strain != "NBC") %>% 
  filter(!MBOA_DMSO_per > 7000) %>% 
  filter(!MBOA_DMSO_per < 0) %>% 
  # filter(Media %in% "MM") %>% 
  ggplot(aes(x= interaction(Strain, AMPO_on_MBOA_plates), y= MBOA_DMSO_per)) +
  geom_boxplot(aes(fill = AMPO_on_MBOA_plates), position=position_dodge()) + 
  geom_jitter(width = 0.1) + 
  # ggpubr::stat_compare_means(label = "p.signif", method = "t.test", ref.group = "DMSO", label.y = 60) +
  # ylim(0, 100) +
  #scale_shape_manual(values = shapes_manual ) +
  # facet_wrap(~Strain, scales = "fixed") +
  scale_fill_manual(values = c("cornflowerblue", "darkred")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0.5)) +
    labs(y = "AUC % MBOA/DMSO",
         x = "",
         title = "Growth of strains in MM",
         fill = "AMPO on plate")
```

Statistics MBOA tolerance in MM

```{r, include=FALSE}
aov_pheno_MM <- results_AUC_tol_MM %>% aov(data = ., MBOA_DMSO_per ~ AMPO_on_MBOA_plates)
summary(aov_pheno_MM) %>% pander()

aov_collection_MM <- results_AUC_tol_MM %>% aov(data = ., MBOA_DMSO_per ~ Collection)
summary(aov_collection_MM) %>% pander()

aov_pheno_strain_MM <- results_AUC_tol_MM %>% aov(data = ., MBOA_DMSO_per ~ AMPO_on_MBOA_plates*Strain)
summary(aov_pheno_strain_MM) %>% pander()

aov_pheno_strain_MM_col <- results_AUC_tol_MM %>% aov(data = ., MBOA_DMSO_per ~ AMPO_on_MBOA_plates*Strain*Collection)
summary(aov_pheno_strain_MM_col) %>% pander()

aov_pheno_strain_col <- results_AUC_tol_MM %>% aov(data = ., MBOA_DMSO_per ~ AMPO_on_MBOA_plates*Collection)
summary(aov_pheno_strain_col) %>% pander()
```

# Phylogenetic tree

```{r, include=FALSE}
results_AUC_mean_MM <- results_AUC %>% filter(Media %in% "MM") %>% group_by(Strain, Chem) %>%
  dplyr::summarise(mean = mean(AUC), n = n()) %>% 
  dplyr::rename(AUC_mean = mean,
                Replicates = n)

results_AUC_mean_MM$Strain <- gsub("NBC", "KHB019", results_AUC_mean_MM$Strain)

results_AUC_mean_MM %<>% mutate(AUC_mean = as.numeric(AUC_mean))
# results_AUC_mean_MBOA <- results_AUC_mean %>% filter(Chem %in% "MBOA") %>% dplyr::select(-Chem)
```

```{r, include=FALSE}
results_AUC_mean_MM_MBOA <- results_AUC_mean_MM %>% filter(Chem %in% "MBOA") 
results_AUC_mean_MM_MBOA_pos <- results_AUC_mean_MM_MBOA
results_AUC_mean_MM_MBOA_pos$AUC_mean <- ifelse(results_AUC_mean_MM_MBOA_pos$AUC_mean < 0, 0, results_AUC_mean_MM_MBOA_pos$AUC_mean)
```

```{r, echo=F, warning=F, message=F}
Fig.S11.MBOA_Micro_carbon <- ggtree::gheatmap(p_host_simple,
         results_AUC_mean_MM_MBOA_pos %>%
         # mutate(AUC = if_else(MBOA_DMSO_per < 0, 0, MBOA_DMSO_per)) %>% 
        # mutate(MBOA_DMSO_per = log10(MBOA_DMSO_per)) %>% 
         # filter(MBOA_DMSO_per < 0) %>% 
         # mutate(Chem = as.character(Conc)) %>%  
         dplyr::select(Strain, AUC_mean) %>% 
         # distinct %>% tidyr::spread(Chem, AUC_mean) %>% 
         as.data.frame() %>% column_to_rownames("Strain") ,
         offset=0.25,
         width=0.1,
         colnames=FALSE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_gradient2(low = "black", mid = "khaki", high ="darkred", midpoint = 1) +
  labs(x = "",
       y = "", 
       shape = "Host plant",
       color = "AMPO phenotype",
       fill = "AUC")

Fig.S11.MBOA_Micro_carbon

# ggsave(plot = Fig.S11.MBOA_Micro_carbon, filename = "Fig.S11.MBOA_Micro_carbon.pdf", width = 15, height = 15, dpi = 300, scale = 1, units = "cm")
# 
# ggsave(plot = Fig.S11.MBOA_Micro_carbon, filename = "Fig.S11.MBOA_Micro_carbon.svg", width = 15, height = 15, dpi = 300, scale = 1, units = "cm")
```

# COMBINED TREE

```{r, echo=F, warning=F, message=F}
met %>% filter(Compound %in% c("MBOA", "HMPAA", "AMPO")) %>% 
  mutate(Compound = factor(Compound, levels = c("MBOA", "HMPAA", "AMPO"))) %>% 
  ggplot(aes(x= Strain, y = uMol)) + 
  geom_bar(stat = "summary") + 
  facet_grid(Treatment~Compound) +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) +
  labs(title = "MBOA metabolites")
```

```{r, echo=F, warning=F, message=F, include=F}
#HMPAA 
met %>% filter(Compound %in% c("HMPAA")) %>%  dplyr::summarize(conc_max = max(uMol)) 
met %>% filter(Compound %in% c("HMPAA")) %>%  filter(Treatment %in% "DMSO") %>% dplyr::summarize(conc_max = max(uMol)) 

met %>% filter(Compound %in% c("HMPAA")) %>%  dplyr::summarize(conc_min = min(uMol))
met %>% filter(Compound %in% c("HMPAA")) %>% filter(uMol > 0) %>% dplyr::summarize(conc_min = min(uMol))
met %>% filter(Compound %in% c("HMPAA")) %>% filter(uMol > 5) %>% dplyr::select(Strain, Treatment)

met %>% filter(Compound %in% c("AMPO")) %>% filter(uMol > 5) %>% dplyr::select(Strain, Treatment)
met %>% filter(Compound %in% c("AMPO")) %>% filter(uMol < 5) %>% dplyr::select(Strain, Treatment)
```


```{r, echo=F, warning=F, message=F, include=F}
metDA %>% 
  filter(Treatment %in% c("G", "D")) %>% 
  filter(Compound %in% c("DIMBOA-Glc", "MBOA", "AMPO")) %>% 
  mutate(Compound = factor(Compound, levels = c("DIMBOA-Glc", "MBOA", "AMPO"))) %>% 
  ggplot(aes(x= Strain, y = uMol)) + 
  geom_bar(stat = "summary") + 
  facet_grid(Treatment~Compound) +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) +
  labs(title = "DIMBOA-Glc metabolites")
```

Thresholds for qualitative analysis
- weak MBOA-degraders: >30% of MBOA degraded compared to the control)
- strong MBOA-degraders: (>90% of MBOA degraded, Fig. 3a). 
strong AMPO-formers (100 - 10 %) 
weak AMPO-formers’, <10% of max. AMPO forming strain)
(0.01 µM, limit of detection).


```{r, echo=F, warning=F, message=F, include=F}
met_MBOA_NBC <- met %>% filter(Strain %in% "NBC") %>% filter(Treatment %in% "MBOA") %>% filter(Compound %in% "MBOA")
met_MBOA_NBC$uMol
# 343.0771 uMol

# 70% 
MBOA_treshold_weak <- (met_MBOA_NBC$uMol / 100) * 70 
MBOA_treshold_weak
# 240.154

# 90 %
MBOA_treshold_strong <- (met_MBOA_NBC$uMol / 100) * 10
MBOA_treshold_strong
# 34.30771

# 0.1 %
MBOA_treshold_LOD <- (met_MBOA_NBC$uMol / 100) * 1
MBOA_treshold_LOD
```

```{r, echo=F, warning=F, message=F, include=F}
met_AMPO_max <- met %>% filter(Treatment %in% "MBOA") %>% filter(Compound %in% c("AMPO")) %>%  dplyr::summarize(conc_max = max(uMol)) 
met_AMPO_max$conc_max
# 471.5786

# 10% of max
AMPO_treshold_strong <- (met_AMPO_max$conc_max / 100) * 10
AMPO_treshold_strong
# 47.15786

# 0.1 % of max
AMPO_treshold_weak <- (met_AMPO_max$conc_max / 100) * 0.1
AMPO_treshold_weak
# 0.4715786
```

```{r, echo=F, warning=F, message=F, include=F}
met_HMPAA_max <- met %>% filter(Treatment %in% "MBOA") %>% filter(Compound %in% c("HMPAA")) %>%  dplyr::summarize(conc_max = max(uMol)) 
met_HMPAA_max$conc_max
# 471.5786

# 10% of max
HMPAA_treshold_strong <- (met_HMPAA_max$conc_max / 100) * 10
HMPAA_treshold_strong
# 4.406674

# 0.1 % of max
HMPAA_treshold_weak <- (met_HMPAA_max$conc_max / 100) * 0.1
HMPAA_treshold_weak
# 0.04406674
```

```{r, echo=F, warning=F, message=F, include=F}
met_DIMBOAGlc_NBC <- metDA %>% filter(Strain %in% "NBC") %>% filter(Treatment %in% "G") %>% filter(Compound %in% "DIMBOA-Glc")
met_DIMBOAGlc_NBC$uMol
# 515.6157

# 70% 
DIMBOAGlc_treshold_weak <- (met_DIMBOAGlc_NBC$uMol / 100) * 70 
DIMBOAGlc_treshold_weak
# 360.931

# 90 %
DIMBOAGlc_treshold_strong <- (met_DIMBOAGlc_NBC$uMol / 100) * 10
DIMBOAGlc_treshold_strong
# 51.56157

# 0.1 %
DIMBOAGlc_treshold_LOD <- (met_DIMBOAGlc_NBC$uMol / 100) * 0.1
DIMBOAGlc_treshold_LOD

# 0.5156157
```

```{r, echo=F, warning=F, message=F, include=F}
metDA_MBOA_max <- met %>% filter(Treatment %in% "MBOA") %>% filter(Compound %in% c("MBOA")) %>%  dplyr::summarize(conc_max = max(uMol)) 
metDA_MBOA_max$conc_max
# 516.2433

# 10% of max
MBOA.DG_treshold_strong <- (metDA_MBOA_max$conc_max / 100) * 10
MBOA.DG_treshold_strong
# 51.62433

# 0.1 % of max
MBOA.DG_treshold_weak <- (metDA_MBOA_max$conc_max / 100) * 0.1
MBOA.DG_treshold_weak
# 0.5162433
```

```{r, echo=F, warning=F, message=F, include=F}
met %>% filter(Strain %in% "Leaf151") %>% filter(Compound %in% c("MBOA", "HMPAA", "AMPO")) %>% filter(Treatment %in% "MBOA")

met %<>% as.data.frame() %>% filter(!RT %in% "NA")
```

```{r, echo=F, warning=F, message=F, include=F}
met <- met %>% as.data.frame() %>% mutate(Compoud = as.factor(Compound))
met$Binary <- NA #add new empty col

# met$uMol[is.na(met$uMol)] = 0

met$Binary[met$Compoud=="HMPAA" & met$uMol<HMPAA_treshold_weak] <- "no" # no HMPAA detected
met$Binary[met$Compoud=="HMPAA" & met$uMol< HMPAA_treshold_strong & met$uMol > HMPAA_treshold_weak] <- "HMPAA_low" #HMPAA not degraded & "MBOA high"
met$Binary[met$Compoud=="HMPAA" & met$uMol> HMPAA_treshold_strong] <- "HMPAA_high" #MBOA not degraded & "MBOA high"

met$Binary[met$Compoud=="AMPO" & met$uMol< AMPO_treshold_weak] <- "no" # no AMPO detected
met$Binary[met$Compoud=="AMPO" & met$uMol<AMPO_treshold_strong & met$uMol>AMPO_treshold_weak+0.001] <- "AMPO_low" #AMPO not degraded & "MBOA high"
met$Binary[met$Compoud=="AMPO" & met$uMol>AMPO_treshold_strong] <- "AMPO_high" #AMPO not degraded & "MBOA high"

met$Binary[met$Compoud=="MBOA" & met$uMol<MBOA_treshold_strong] <- "no" # no MBOA detected
met$Binary[met$Compoud=="MBOA" & met$uMol<MBOA_treshold_weak & met$uMol>MBOA_treshold_strong] <- "MBOA_low" #MBOA not degraded & "MBOA high"
met$Binary[met$Compoud=="MBOA" & met$uMol>MBOA_treshold_weak] <- "MBOA_high" #MBOA not degraded & "MBOA high"

metDA <- metDA %>% filter(Treatment %in% "G") %>% as.data.frame() %>% mutate(Compoud = as.factor(Compound))
metDA$Binary <- NA #add new empty col

metDA$Binary[metDA$Compoud=="DIMBOA-Glc" & metDA$uMol<DIMBOAGlc_treshold_strong] <- "no" # no MBOA detected
metDA$Binary[metDA$Compoud=="DIMBOA-Glc" & metDA$uMol<DIMBOAGlc_treshold_weak & metDA$uMol>DIMBOAGlc_treshold_strong] <- "DIMBOA-Glc_low" #MBOA not degraded & "MBOA high"
metDA$Binary[metDA$Compoud=="DIMBOA-Glc" & metDA$uMol>DIMBOAGlc_treshold_weak] <- "DIMBOA-Glc_high" #MBOA not degraded & "MBOA high"

metDA$Binary[metDA$Compoud=="MBOA" & metDA$uMol<MBOA.DG_treshold_weak] <- "no" # no MBOA detected
metDA$Binary[metDA$Compoud=="MBOA" & metDA$uMol<MBOA.DG_treshold_strong & metDA$uMol>MBOA.DG_treshold_weak] <- "MBOA_low" #MBOA not degraded & "MBOA high"
metDA$Binary[metDA$Compoud=="MBOA" & metDA$uMol>MBOA.DG_treshold_strong] <- "MBOA_high" #MBOA not degraded & "MBOA high"
```

```{r, echo=F, warning=F, message=F, include=F}
results_AUC_mean_MM_MBOA_pos %>% ggplot(aes(x = Strain, y = AUC_mean)) + geom_bar(stat = "summary")
```

```{r, echo=F, warning=F, message=F, include=F}
results_AUC_mean_MM_MBOA %>% filter(Strain %in% c("LWH11", "Leaf320"))

results_AUC_mean_MM_MBOA_max <- results_AUC_mean_MM_MBOA %>% as.data.frame() %>%  dplyr::summarize(conc_max = max(AUC_mean)) 
results_AUC_mean_MM_MBOA_max$conc_max
# 516.2433

# 10% of max
results_AUC_mean_MM_MBOA_max_treshold <- (results_AUC_mean_MM_MBOA_max$conc_max / 100) * 20

results_AUC_mean_MM_MBOA %<>% mutate(Binary = case_when(AUC_mean < results_AUC_mean_MM_MBOA_max_treshold ~ "no growth",
                                                           AUC_mean > results_AUC_mean_MM_MBOA_max_treshold ~ "growth"))
```


Tree with AMPO Phenotype on plates

```{r, warning=F, echo=F, message=F}
tree_plate <- ggtree::gheatmap(p_host_simple + new_scale_fill(),
         metadata_tree %>% 
         dplyr::select(Strain, AMPO_on_MBOA_plates) %>% 
           dplyr::rename('Plate assay' = AMPO_on_MBOA_plates) %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain"),
         offset=0.2,
         width=0.1,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("lightgoldenrod", "darkred")) +
  labs(fill = "AMPO phenotype")

```

Tree as above with MBOA metabolites

```{r, echo=F, warning=F, message=F}
tree_plate_MBOA <- ggtree::gheatmap(tree_plate + new_scale_fill(),
         met %>% filter(Compound %in% c("MBOA", "HMPAA", "AMPO")) %>% 
         filter(Treatment %in% "MBOA") %>% 
         mutate(Compound = factor(Compound, levels = c("MBOA", "AMPO", "HMPAA"))) %>% 
         # mutate(Binary = factor(Binary, levels = c("MBOA high", "MBOA low", "AMPO high", "AMPO low", "HMPAA high", "HMPAA low", "not detected"))) %>% 
         # filter(!Binary %in% c("AMPO low", "HMPAA low")) %>% 
         dplyr::select(Strain, Compound, Binary) %>% unique() %>% 
         dplyr::select(Strain, Compound, Binary) %>% tidyr::spread(Compound, Binary) %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain") ,
         offset=0.25,
         width=0.3,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c( "darkred", "white", "gold2", "white", "darkblue", "white"), na.value = "white") +
  # labels = c( "AMPO high", "AMPO low", "MBOA high", "MBOA low", "HMPAA high", "HMPAA low", "not detected"), +
  labs(#x = "Benzoxazinoid",
       #y = "Relative AUC", 
       colour = "AMPO phenotype",
       shape = "Host plant",
       fill = "Concentration") 

#dev.off()
```

Tree as above with DIMBOA-Glc metabolites 

```{r, echo=F, warning=F, message=F}
tree_plate_MBOA_DMG <- ggtree::gheatmap(tree_plate_MBOA + new_scale_fill(),
         metDA %>% filter(Treatment %in% c("G")) %>% 
           filter(Compound %in% c("DIMBOA-Glc", "MBOA")) %>% 
           filter(!uMol %in% 0.00000000) %>%
           filter(!uMol > 600) %>%
           mutate(Compound = factor(Compound, levels = c("DIMBOA-Glc", "MBOA"))) %>% 
         dplyr::select(Strain, Compound, Binary) %>% tidyr::spread(Compound, Binary) %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain") ,
         offset=0.37,
         width=0.2,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("darkorange2",  "orange", "darkblue", "#555CCB", "white"), na.value = "white")+
  labs(#x = "Benzoxazinoid",
       #y = "Relative AUC", 
       colour = "AMPO phenotype",
       shape = "Host plant",
       fill = "Concentration") 


```

```{r, echo=F, warning=F, message=F}
tree_plate_MBOA_DMG_carbon <- ggtree::gheatmap(tree_plate_MBOA_DMG + new_scale_fill(),
                                        results_AUC_mean_MM_MBOA %>%
         dplyr::select(Strain, Binary) %>% 
         dplyr::rename('C-source' = Binary) %>% 
         as.data.frame() %>% column_to_rownames("Strain") ,
         offset=0.46,
         width=0.1,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         # colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("khaki4",  "white"))+
  labs(x = "",
       y = "", 
       shape = "Host plant",
       color = "AMPO phenotype",
       fill = "AUC")


```

```{r, echo=F, warning=F, message=F}
tree_plate_MBOA_DMG_carbon_OG <- ggtree::gheatmap(tree_plate_MBOA_DMG_carbon + new_scale_fill(),
         met_OG_selected %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain"),
         offset=0.52,
         width=0.5,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("gray100", "gray80", "gray60", "gray40", "gray20", "white")) +
  labs(fill = "Orthogroup copies")

tree_plate_MBOA_DMG_carbon_OG

# ggsave(plot = tree_plate_MBOA_DMG_carbon_OG, filename = "tree_plate_MBOA_DMG_carbon_OG.pdf", width = 23, height = 20, dpi = 300, scale = 1, units = "cm")
# # 
# ggsave(plot = tree_plate_MBOA_DMG_carbon_OG, filename = "tree_plate_MBOA_DMG_carbon_OG.svg", width = 26, height = 20, dpi = 300, scale = 1, units = "cm")
```

```{r, warning=F, echo=F, message=F}
sessionInfo()
```

