---
title: "Metabolite screen maize root bacteria"
subtitle: "Thoenen et al. AMPO formation"
author: "Lisa Thönen"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_caption: yes
    toc: true
    toc_float: true
    theme: cerulean
    code_folding: show
editor_options: 
  chunk_output_type: console
---


Here we analyse the metabolite data from the maize root bacteria screening experiment in December 2019. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE}
rm(list=ls())
```

```{r,include=FALSE}
# install.packages("tidyverse")
# install.packages("magrittr")
# install.packages("readxl")
# install.packages("pheatmap")
# install.packages("reshape2")
# install.packages("grid")
```

```{r,include=FALSE}
library("tidyverse")
library("magrittr")
library("readxl")
library("ggpmisc")
library("pander")
library("RColorBrewer")
library("pheatmap")
library("reshape2")
library("grid")

# load functions
source("Input/functions/vennDia.R")
source("Input/functions/fun_venn_plot.R")
source("Input/functions/fun_venn_pair.R")
source("Input/functions/fun_venn_triple.R")
```

```{r, warning=F, echo=F, message=F}
metadata <- read.csv("Input/metadata60strains.csv", sep = ";")

database <- read_excel("Input/Database_MRB_isolates.xlsx")

metadata$Genus <- gsub("Allorhizobium-Neorhizobium-Pararhizobium-Rhizobium", "Rhizobium", metadata$Genus) 

metadata$Genus[is.na(metadata$Genus)] = "no"
metadata$Phylum[is.na(metadata$Phylum)] = "no"
metadata$Genus <- gsub("Microbacterium", "Micrococcineae", metadata$Genus)
database$Genus <- gsub("Moraxellaceae", "Acinetobacter", database$Genus)


# unified taxonomy with chapter 1
metadata <- left_join(metadata %>% dplyr::select(Strain, Nr, Plate), database, by = "Strain") %>% dplyr::select(Strain, strain_old, Nr, Plate, Phylum, Class, Order, Family, Genus)

metadata$Family <- gsub("Pseudomonadales", "Moraxellaceae", metadata$Family)
```

```{r,include=FALSE}
BXData_all <- read.table("Input/20191209_BacBXD_quantified.txt", header = TRUE)
BXData_all %>% colnames()


# BXData$uMol_log <- log10(BXData$uMol)*-1

BXData_all$uMol[is.na(BXData_all$uMol)] = 0
BXData_all$Area[is.na(BXData_all$Area)] = 0
BXData_all$ng.mL_QuanLynx[is.na(BXData_all$ng.mL_QuanLynx)] = 0
BXData_all$ng.mL_adj[is.na(BXData_all$ng.mL_adj)] = 0

BXData_all$ng.mL_adj[is.na(BXData_all$ng.mL_adj)] = 0
# BXData_all$uMol_log[is.na(BXData_all$uMol_log)] = 0

# unify strain names across data tables
BXData_all$Strain <- gsub("LRC7-S", "LRC7.S", BXData_all$Strain)
BXData_all$Strain <- gsub("LRC7-O", "LRC7.O", BXData_all$Strain)
BXData_all$Strain <- gsub("LRH8-S", "LRH8.S", BXData_all$Strain)
BXData_all$Strain <- gsub("LRH8-O", "LRH8.O", BXData_all$Strain)
BXData_all$Strain <- gsub("LPB4-R", "LPB4.R", BXData_all$Strain)
BXData_all$Strain <- gsub("LPB4-O", "LPB4.O", BXData_all$Strain)

# filter out the LMI1 samples which were replicated in each qTOF run
BXData_all <- BXData_all %>% 
  filter(Name != "20191210_LT_NBC_MBOA") %>%
  filter(Name != "20191210_LT_NBC_DIMBOA") %>%
  filter(Name != "20191210_LT_NBC_DG") %>%
  filter(Name != "20191210_LT_NBC_Ctrl") %>%
  filter(Name != "20191210_LT_LMI1_MBOA_02") %>% 
  filter(Name != "20191210_LT_LMI1_MBOA") %>% 
  filter(Name != "20191211_LT_LMI1_MBOA") 
```

Some strains had to be excluded because they grew bad in the assay. This was checked by optical density measurement of the 96 well plates (which was prone to give false results because there condensed water forms below the lid which affects the measurements) and by measuring two amino acids in the cultures. LMR1 was excluded because of unsure taxonomy (Paenib. or neorhizobium). The antibiotic resistant strains LPB4.R, "LRC7.S and LRH8.S were also excluded and only their orignals (xyz.O) kept.

```{r, warning=F, echo=F, message=F}
metadata %>% filter(Strain %in% c("LRH8.O", "LST17", "LMC1", "LMK1", "LME2", "LMX8", "LMX3", "LMX11", "LMR1", "LPB4.R", "LRC7.S", "LRH8.S")) %>% dplyr::select(Strain, Genus, Phylum) %>% knitr::kable()
```

```{r, warning=F, echo=F, message=F}
BXData_all %<>% filter(!Strain %in% c("LRH8.O", "LST17", "LMC1", "LMK1", "LME2", "LMX8", "LMX3", "LMX11", "LMR1", "LPB4.R", "LRC7.S", "LRH8.S"))
```

```{r,include=FALSE}
# Create separate tables for all the treatments & combine BXData table with metadata
# filter out strains with strange growth data

# MBOA
BXData <- BXData_all %>% filter(Treatment %in% "MBOA")
BXData <- left_join(BXData, metadata, by = "Strain")
BXData$Genus_Strain <- as.factor(paste(BXData$Genus, BXData$Strain, sep="_") )

# DIMBOA
BXData_DIMBOA <- BXData_all %>% filter(Treatment %in% "DIMBOA")
BXData_DIMBOA <- left_join(BXData_DIMBOA, metadata, by = "Strain")
BXData_DIMBOA$Genus_Strain <- as.factor(paste(BXData_DIMBOA$Genus, BXData_DIMBOA$Strain, sep="_") )

# DIMBOA-Glc
BXData_DG <- BXData_all %>% filter(Treatment %in% "DG")
BXData_DG <- left_join(BXData_DG, metadata, by = "Strain")
BXData_DG$Genus_Strain <- as.factor(paste(BXData_DG$Genus, BXData_DG$Strain, sep="_") )
```

```{r, warning=F, echo=F, message=F, include=F}
# extract the NBC data for each compound (later used for calculation of % compound metabolized)
MBOA_NBC <- BXData_all %>% filter(Strain %in% "NBC", Treatment %in% "MBOA", Compound_name %in% "MBOA", Block %in% "1")
MBOA_NBC$uMol

DIMBOA_NBC <- BXData_all %>% filter(Strain %in% "NBC", Treatment %in% "DIMBOA", Compound_name %in% "DIMBOA", Block %in% "1")
DIMBOA_NBC$uMol

DIMBOA_NBC_MBOA <- BXData_all %>% filter(Strain %in% "NBC", Treatment %in% "DIMBOA", Compound_name %in% "MBOA", Block %in% "1")
DIMBOA_NBC_MBOA$uMol

DIMBOAGlc_NBC <- BXData_all %>% filter(Strain %in% "NBC", Treatment %in% "DG", Compound_name %in% "DIMBOA-Glc", Block %in% "1")
DIMBOAGlc_NBC$uMol
```

**Dilutions**

All the samples (except T0 samples) were diluted with acidified MeOH in a ratio (30:70 - sample:MeOH). Then they were diluted 1:6 in the pooled eppi and 1:10 in the analysis vial, resulting in a final dilution of 200. The T0 samples were diluted 1:6. Therefore the values have to be adjusted to the dilution.

```{r, warning=F, echo=F, message=F, include=F}
BXData_all %<>%
  mutate(uMol = case_when(Strain != "T0" & uMol != 0  ~ uMol*200,
                          Strain == "T0" ~ uMol))

BXData_all %<>%
  mutate(uMol = case_when(Strain == "T0" & uMol != 0  ~ uMol*6,
                          Strain != "T0" ~ uMol))

BXData_all %>% filter(Strain %in% "NBC", Compound_name %in% "MBOA")
 
BXData_all %>% filter(Strain %in% "T0", Compound_name %in% "MBOA")
 
BXData %<>%
  mutate(uMol = case_when(Strain != "T0" & uMol != 0  ~ uMol*200,
                          Strain == "T0" ~ uMol))

BXData %<>%
  mutate(uMol = case_when(Strain == "T0" & uMol != 0  ~ uMol*6,
                          Strain != "T0" ~ uMol))

BXData_DIMBOA %<>%
  mutate(uMol = case_when(Strain != "T0" & uMol != 0  ~ uMol*200,
                          Strain == "T0" ~ uMol))

BXData_DIMBOA %<>%
  mutate(uMol = case_when(Strain == "T0" & uMol != 0  ~ uMol*6,
                          Strain != "T0" ~ uMol))

BXData_DG %<>%
  mutate(uMol = case_when(Strain != "T0" & uMol != 0  ~ uMol*200,
                          Strain == "T0" ~ uMol))

BXData_DG %<>%
  mutate(uMol = case_when(Strain == "T0" & uMol != 0  ~ uMol*6,
                          Strain != "T0" ~ uMol))
```

# BX profiles of media controls
Benzoxazinoids were mixed to TSB growth media, each at a concentration of 500 uM. BXs were MBOA, DIMBOA and DIMBOA-Glc. 
The treatments were diluted and frozen directly at the start of the experiment (T0) and incubated without bacteria throughout the experiment (NBC). The T0 samples were differently diluted at the beginning, therefore they were here normalized with MBOA values (*35) - MBOA is stable in these conditions. 
NBC and T0 plotted show the spontaneous degradation of the BXs in the culture growth conditions. 

MBOA seems to be pure and stable over the course of the experiment. DIMBOA is not stable and degrades completly to MBOA. As the profile shows at T0 it is not pure and also contains HMBOA which degrades to BOA. 

DIMBOA-Glc is stable but not pure. 

```{r, warning=F, echo=F, message=F}
BXData_all %>% mutate(Compound_name = as.factor(Compound_name)) %>% mutate(Strain = as.factor(Strain)) %>% 
  filter(Strain %in% c("T0", "NBC")) %>%
  ggplot(aes(x = Compound_name, y = uMol)) + 
  geom_bar(aes(fill = Strain), position = "dodge", stat = "summary", show.legend = TRUE) +
  facet_grid(~Treatment, scales = "free_y") +
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0.5 )) +
  labs(x = "",
       fill = "")
```

```{r, warning=F, echo=F, message=F}
media.To <- BXData_all %>% mutate(Compound_name = as.factor(Compound_name)) %>% mutate(Strain = as.factor(Strain)) %>% 
  filter(!Treatment %in% "DIMBOA") %>% 
  filter(Compound_name %in% c("MBOA", "HMBOA-Glc", "HMBOA", "DIMBOA-Glc", "DIMBOA", "DIM2BOA-Glc", "BOA")) %>% 
  filter(Strain %in% c("T0", "NBC")) %>%
  mutate(Treatment = gsub("DG", "DIMBOA-Glc", Treatment)) %>% 
  mutate(Treatment = gsub("Ctrl", "DMSO", Treatment)) %>% 
  mutate(Treatment = factor(Treatment, levels = c("MBOA", "DIMBOA-Glc", "DMSO"))) %>% 
  ggplot(aes(x = Compound_name, y = uMol)) + 
  geom_bar(aes(fill = Strain), position = "dodge", stat = "summary", show.legend = TRUE) +
  facet_grid(~Treatment, scales = "free_y") +
  theme_bw() +
  theme(strip.background = element_rect(color = NULL, fill= "white", linetype="solid"), strip.text.x = element_text(size = 16/.pt, margin = margin(1, 1, 1, 1)))+
  theme(axis.text.x=element_text(size = 16/.pt, angle = -90, hjust = 0, vjust = 0.5),
        axis.text.y=element_text(size = 16/.pt),
        axis.title = element_text(size = 16/.pt)) +
  scale_fill_manual(values = c("grey20", "grey60"), labels = c("end", "start"))+
  coord_flip()+
  labs(x = "",
       y = "Concentration [uM]",
       fill = "")

media.To

ggsave(plot = media.To, filename = "FigS5b.svg", width = 9, height = 4, dpi = 300, scale = 1, units = "cm")
```


# MBOA assay

## Qualitative analysis

Since there is quite some variation in the absolute values of metabolites in the cultures, which makes it difficult to categorize the bacteria to metabolic types, we to qualitative categorization of the bacteria to groups. We define threshold concentrations of the supplemented compounds (MBOA & DIMBOA-Glc) to group bacteria into "degraders" or "non-degraders" and for metabolization products (AMPO & AAMPO) to group in "formers" and "non-formers". 

```{r, warning=F, echo=F, message=F}
# All strains
BXData_all %<>% filter(!Strain %in% c("Standard", "T0"))
All <- BXData_all$Strain %>% unique()
```

### classification

**MBOA degradation**

We define thresholds for *strong*, *weak* and *non-degraders* of MBOA relative to the MBOA levels detected in the NBC sample (587.5834 μM):  

*strong MBOA degraders*: ≤10% of MBOA levels compared to NBC detected (10% = 58.75834).   

*weak MBOA degraders*: 70% > x > 10% of MBOA levels compared to NBC detected.   

*non-degraders of MBOA*: ≥70% of MBOA levels compared to NBC detected (70% = 411.3084).  

```{r, warning=F, echo=F, message=F, include=F}
## data.frame with MBOA levels of each strain
# MBOA <- BXData_all %>% filter(Treatment %in% "MBOA") %>% filter(Compound_name %in% "MBOA") %>% select(Strain, uMol) %>% unique() %>% na.omit()

## exploring MBOA levels to define a threshold for MBOA metabolization
# MBOA %<>% dplyr::arrange(desc(uMol))
# MBOA$uMol %>% max()
# MBOA$uMol %>% min()

## exploring MBOA levels of different strains
# MBOA %>% filter(Strain %in% "LMI1x") 
# MBOA %>% filter(Strain %in% "LME3") 
# MBOA %>% filter(Strain %in% "LBA112") 
# MBOA %>% filter(Strain %in% "LAC11") 

## MBOA levels in NBC, no bacteria control sample
MBOA_NBC_uMol <- BXData_all %>%
  filter(Strain != "T0") %>% 
  filter(Compound_name %in% "MBOA") %>% 
  filter(Treatment %in% "MBOA") %>% 
  filter(Strain %in% "NBC") %>% 
  dplyr::select(Strain, Compound_name, uMol) %>% 
  arrange(desc(uMol)) %>% 
  select(uMol)

### define thresholds for strong, weak and non-degraders of MBOA
## v1 threshold relative to NBC sample
## weak is ≤70% NBC
MBOA_met_threshold_weak <- (MBOA_NBC_uMol$uMol / 100) * 70   # 411.3084
# MBOA_met_threshold <- "400"   # option to set manually to 400
## strong is ≤10% NBC (90% degraded)
MBOA_met_threshold_strong <- (MBOA_NBC_uMol$uMol / 100) * 10 # 58.75834

## v2 treshold is relative to 500 µM input
## weak is ≤70% NBC
# MBOA_met_threshold_weak <- 350
## strong is ≤10% NBC (90% degraded)
# MBOA_met_threshold_strong <- 50
### we decided to go for v1, more stringent

```

**AMPO formation**

We define thresholds for *non-*, *weak* or *strong* AMPO formers relative to maximum level of AMPO formed by one of the strains (42.49294 µM):  

*non-formers of AMPO:* form <0.1% of AMPO compared to the max. AMPO-former (0.1% = 0.04249294 µM).  

*weak AMPO-formers:* form 10% > x 0.1% of AMPO compared to the max. AMPO-former.  

*Strong AMPO-formers:* form >10% of AMPO compared to the max. AMPO-former (10% = 4.249294 µM).  

```{r, warning=F, echo=F, message=F, include=F}

## define max level of AMPO formed by one of the strains
met_AMPO_from_MBOA_max <- BXData_all %>% filter(Treatment %in% "MBOA") %>% 
                                         filter(Compound_name %in% c("AMPO")) %>%  
                                         filter(!uMol %in% NA) %>%  
                                         dplyr::summarize(conc_max = max(uMol))

AMPO_from_MBOA_strong <- met_AMPO_from_MBOA_max$conc_max / 100 * 10
# 4.249294

AMPO_from_MBOA_weak <- met_AMPO_from_MBOA_max$conc_max / 100 * 0.1
# 0.04249294

```

**AAMPO formation**

We did not define thresholds for AAMPO formers, but recorded if AAMPO was detected in a culture.  

```{r, warning=F, echo=F, message=F, include=F}
# MBOA_AAMPO <- BXData_all %>% filter(Treatment %in% "MBOA") %>% filter(Compound_name %in% "AAMPO", uMol > 0.13) %>% dplyr::select(Strain, uMol) %>% na.omit() %>% unique() %>% as.data.frame()
# 
# MBOA_AAMPO <- MBOA_AAMPO$Strain
# MBOA_AAMPO %>% pander()
```

```{r, warning=F, echo=F, message=F}
# Prepare data
# make long
BXData_long_mol <- BXData %>% filter(Strain != "T0") %>% 
  filter(Strain != "Standard") %>%  
  filter(Treatment %in% "MBOA") %>% 
  filter(Name != "20191210_LT_LMI1_MBOA_02") %>% 
  filter(Name != "20191211_LT_LMI1_MBOA") %>%  
  filter(Name != "20191210_LT_NBC_MBOA") %>%
  filter(Treatment %in% "MBOA") %>% 
  filter(Compound_name %in% c("MBOA", "AMPO", "AAMPO")) %>% 
  mutate(Genus = gsub( "Acinetobacter", "Moraxellaceae",Genus)) %>% 
  arrange(desc(Family)) %>% 
  mutate(Genus_Strain = gsub("NA_NBC", "_NBC", Genus_Strain)) %>%  
  dplyr::select(Genus_Strain, Compound_name, uMol) %>%  
  unique() %>% 
  tidyr::spread(Compound_name, uMol) %>% 
  column_to_rownames("Genus_Strain")

BXData_long_mol.matrix <- as.matrix(BXData_long_mol)
BXData_long_mol.matrix[is.na(BXData_long_mol.matrix)] = 0

# Change column order
BXcols <- c("MBOA", "AMPO", "AAMPO")

BXData_long_mol.matrix <- BXData_long_mol.matrix[, BXcols]

### translate thresholds to classifications (numeric for color code)

## MBOA degraders
# strong MBOA degraders / no MBOA detected
BXData_long_mol.matrix[, 1][BXData_long_mol.matrix[, 1]>=0 & BXData_long_mol.matrix[, 1]<=MBOA_met_threshold_strong] <- 0
# weak MBOA degraders / MBOA degraded
BXData_long_mol.matrix[, 1][BXData_long_mol.matrix[, 1]>=MBOA_met_threshold_strong & BXData_long_mol.matrix[, 1]<=MBOA_met_threshold_weak] <- 1 
# non-degraders / MBOA not degraded
BXData_long_mol.matrix[, 1][BXData_long_mol.matrix[, 1]>=MBOA_met_threshold_weak & BXData_long_mol.matrix[, 1]<=700] <- 2

### Reclassification
# This assay here would classify the strains LBA112 and LMI1x as *weak* MBOA-degraders. However, two replicate experiments (Fig. 2c, Fig. S11a) revealed that they do not degrade MBOA. Hence the classifications of these two strains were manually adjusted to *non-degraders*. 
BXData_long_mol.matrix["Bacillus_LBA112",1] <- 2
BXData_long_mol.matrix["Micrococcineae_LMI1x",1] <- 2

## AMPO formers
# non-formers / no AMPO detected
BXData_long_mol.matrix[, 2][BXData_long_mol.matrix[, 2]>=0 & BXData_long_mol.matrix[, 2]<=AMPO_from_MBOA_weak] <- 0 
# weak AMPO formers / little AMPO detected
BXData_long_mol.matrix[, 2][BXData_long_mol.matrix[, 2]>=AMPO_from_MBOA_weak & BXData_long_mol.matrix[, 2]<=AMPO_from_MBOA_strong] <- 3
# strong AMPO formers / lots of AMPO found
BXData_long_mol.matrix[, 2][BXData_long_mol.matrix[, 2]>=AMPO_from_MBOA_strong & BXData_long_mol.matrix[, 2]<=100] <- 4 

## AAMPO detected
BXData_long_mol.matrix[, 3][BXData_long_mol.matrix[, 3]>=0 & BXData_long_mol.matrix[, 3]<=AMPO_from_MBOA_weak] <- 0 # no AAMPO detected
BXData_long_mol.matrix[, 3][BXData_long_mol.matrix[, 3]>=AMPO_from_MBOA_weak] <- 5 

heatmapMBOAFig.3_binary <- pheatmap(BXData_long_mol.matrix, 
                             color = c("white", "#555CCB", "darkblue", "tomato1", "darkred", "firebrick1"),
                             cellwidth = 10, 
                             cutree_rows = 6, 
                             fontsize = 8, 
                             cluster_cols = FALSE, 
                             cluster_rows = FALSE, 
                             main = "MBOA")


ggsave(plot = heatmapMBOAFig.3_binary,  filename = "Fig2a.svg", width = 14, height = 21, dpi = 300, scale = 1, units = "cm")
```


## Quantitative analysis

```{r, warning=F, echo=F, message=F}
BXData_all$cols <- as.character(BXData_all$Compound_name)
BXData_all[BXData_all$Compound_name == "AMPO", ]$cols <- "darkred"
BXData_all[BXData_all$Compound_name == "AAMPO", ]$cols <- "firebrick1"
# BXData_all[BXData_all$Compound_name == "MHPA", ]$cols <- "gold2"
BXData_all[BXData_all$Compound_name == "MBOA", ]$cols <- "darkblue"

BXData_all[BXData_all$Compound_name == "DIM2BOA-Glc", ]$cols <-  "sandybrown"
BXData_all[BXData_all$Compound_name == "DIMBOA-Glc", ]$cols <-  "darkorange2"
BXData_all[BXData_all$Compound_name == "DIMBOA", ]$cols <-  "tan"
BXData_all[BXData_all$Compound_name == "BOA", ]$cols <- "slateblue"

BXData_all[BXData_all$Compound_name == "APO", ]$cols <-  "pink"
BXData_all[BXData_all$Compound_name == "AAPO", ]$cols <-  "pink3"

BXData_all[BXData_all$Compound_name == "HDM2BOA-Glc", ]$cols <-  "thistle1"
BXData_all[BXData_all$Compound_name == "HDMBOA-Glc", ]$cols <-  "thistle3"

BXData_all[BXData_all$Compound_name == "HM2BOA-Glc", ]$cols <-  "skyblue4"
BXData_all[BXData_all$Compound_name == "HMBOA-Glc", ]$cols <-  "skyblue2"
BXData_all[BXData_all$Compound_name == "HMBOA", ]$cols <- "skyblue"

temp <- data.frame(BXData_all$Compound_name, BXData_all$cols)
temp <- plyr::ddply(temp, .variables="BXData_all.Compound_name", .fun=unique)   #library(plyr)
level_cols_Compound <- as.character(temp[,2])
names(level_cols_Compound) <- temp[,1]
```

```{r, warning=F, echo=F, message=F, include=F}
BXData_all_tax <- left_join(BXData_all, metadata %>% dplyr::select(Strain, Genus, Family, Phylum), by = "Strain")
```

### bargraph all compounds

```{r, warning=F, echo=F, message=F, include=T}
MBOA.bargraph <- BXData_all_tax %>% 
  mutate(Strain_Genus = interaction(Strain, Genus)) %>% 
  filter(Strain != "T0") %>% 
  filter(Treatment %in% c("MBOA")) %>% 
  mutate(Treatment = gsub("MBOA", "DIMBOA-Glc", Treatment)) %>%  ################################# what is this?
  filter(Strain != "Standard") %>% 
  ggplot(aes(y = uMol, x = fct_rev(Strain_Genus))) + 
  geom_bar(aes(fill = Compound_name), position = "stack", stat = "summary", fun.y ="mean") +
  scale_fill_manual(values = level_cols_Compound) +
  theme_bw() +
  scale_y_continuous(expand = expansion(mult = c(0, 0)))+
  theme(strip.background = element_rect(color = NULL, fill= "white", linetype="solid"), strip.text.x = element_text(size = 16/.pt, margin = margin(1, 1, 1, 1)))+
  theme(axis.text.x=element_text(size = 16/.pt, angle = -90, hjust = 0, vjust = 0.5),
        axis.text.y=element_text(size = 16/.pt),
        axis.title = element_text(size = 16/.pt)) +
  coord_flip()+
  labs(x = "",
       y = "Concentration [µM]",
       fill = "") +
  geom_hline(yintercept=MBOA_met_threshold_weak, color = "dimgrey")+ 
  geom_hline(yintercept=MBOA_met_threshold_strong, color = "dimgrey") 

MBOA.bargraph

ggsave(plot = MBOA.bargraph, filename = "FigS5a_MBOAbargraph.svg", width = 10, height = 12, dpi = 300, scale = 1, units = "cm")
```

### bargraph (A)AMPO 

```{r, warning=F, echo=F, message=F}
AMPO.MBOA.bargraph <- BXData_all_tax %>% 
  mutate(Strain_Genus = interaction(Strain, Genus)) %>% 
  filter(Strain != "T0") %>% 
  filter(Compound_name %in% c("AMPO", "AAMPO")) %>% 
  filter(Treatment %in% c("MBOA")) %>% 
  mutate(Treatment = gsub("MBOA", "DIMBOA-Glc", Treatment)) %>% 
  filter(Strain != "Standard") %>% 
  ggplot(aes(y = uMol, x = fct_rev(Strain_Genus))) + 
  geom_bar(aes(fill = Compound_name), position = "stack", stat = "summary", fun.y ="mean") +
  facet_wrap(~Compound, ncol=2, scales="free_x") +
  scale_fill_manual(values = level_cols_Compound) +
  theme_bw() +
  scale_y_sqrt(expand=expansion(mult = c(0, 0.1))) +
  theme(strip.background = element_rect(color = NULL, fill= "white", linetype="solid"), strip.text.x = element_text(size = 16/.pt, margin = margin(1, 1, 1, 1)))+
  theme(axis.text.x=element_text(size = 16/.pt, angle = -90, hjust = 0, vjust = 0.5),
        axis.text.y=element_text(size = 16/.pt),
        axis.title = element_text(size = 16/.pt)) +
  coord_flip()+
  labs(x = "", y = "Concentration [µM]", fill = "") +
  geom_hline(yintercept = AMPO_from_MBOA_strong, color = "dimgrey")

AMPO.MBOA.bargraph

ggsave(plot = AMPO.MBOA.bargraph, filename = "FigS5a_AMPO_MBOAbargraph.svg", width = 9, height = 12, dpi = 300, scale = 1, units = "cm")
```



# DIMBOA-Glc assay

## Qualitative analysis

### classification

**DIMBOA-Glc degradation**

We define thresholds for *strong*, *weak* and *non-degraders* of DIMBOA-Glc relative to the DIMBOA-Glc levels detected in the NBC sample (246.9373 μM). Of note, this DIMBOA-Glc amounts were lower compared to MBOA because the DIMBOA-Glc we used, was of 70% purity. 

*strong DIMBOA-Glc degraders*: ≤10% of DIMBOA-Glc levels compared to NBC detected (10% = 24.69373 µM).   

*weak DIMBOA-Glc degraders*: 70% > x > 10% of DIMBOA-Glc levels compared to NBC detected.   

*non-degraders of DIMBOA-Glc*: ≥70% of DIMBOA-Glc levels compared to NBC detected (70% = 172.8561 µM).  

```{r, warning=F, echo=F, message=F, include=F}
# ## data.frame with  DIMBOA-Glucose levels of each strain
# DG <- BXData_all %>% filter(Treatment %in% "DG") %>% filter(Compound_name %in% "DIMBOA-Glc") %>% select(Strain, uMol) %>% unique() 
# 
# ## exploring DG levels to define a threshold for DG metabolization
# DG %<>% dplyr::arrange(desc(uMol))
# DG$uMol %>% max()
# DG$uMol %>% min()
# 
# ## exploring DG levels of different strains
# DG %>% filter(Strain %in% "LME3") # NA
# DG %>% filter(Strain %in% "NBC")  # 246.9373

## DG levels in NBC, no bacteria control sample
DG_NBC_uMol <- BXData_all %>%
  filter(Strain != "T0") %>% 
  filter(Compound_name %in% "DIMBOA-Glc") %>% 
  filter(Treatment %in% "DG") %>% 
  filter(Strain %in% "NBC") %>% 
  dplyr::select(Strain, Compound_name, uMol) %>% 
  arrange(desc(uMol)) %>% 
  select(uMol)

### define thresholds for strong, weak and non-degraders of DG
## threshold relative to NBC sample
## weak is ≤70% NBC
DMG_met_threshold_weak <- (DG_NBC_uMol$uMol / 100) * 70   # 172.8561
## strong is ≤10% NBC (90% degraded)
DMG_met_threshold_strong <- (DG_NBC_uMol$uMol / 100) * 10 # 24.69373

```

**MBOA formation after DIMBOA-Glc degradation**

We define thresholds for *non-*, *weak* or *strong* MBOA formers relative to maximum level of MBOA formed by one of the strains in DIMBOA-Glc medium (143.2905 µM):  

*non-formers of MBOA:* form <0.1% of MBOA compared to the max. MBOA-former (0.1% = 0.1432905 µM).  

*weak MBOA-formers:* form 10% > x 0.1% of MBOA compared to the max. MBOA-former.  

*Strong MBOA-formers:* form >10% of MBOA compared to the max. MBOA-former (10% = 14.32905 µM).  

```{r, warning=F, echo=F, message=F, include=F}
met_MBOA_from_DMG_max <- BXData_all %>% filter(Treatment %in% "DG") %>% 
                                         filter(Compound_name %in% c("MBOA")) %>%  
                                         filter(!uMol %in% NA) %>%  
                                         dplyr::summarize(conc_max = max(uMol))

MBOA_from_DMG_strong <- met_MBOA_from_DMG_max$conc_max / 100 * 10
# 14.32905

MBOA_from_DMG_weak <- met_MBOA_from_DMG_max$conc_max / 100 * 0.1
# 0.1432905
```

**AMPO formation after DIMBOA-Glc degradation**

We took the same thresholds for *non-*, *weak* or *strong* AMPO formers for the DIMBOA-Glc assay as in the MBOA assay.  

```{r, warning=F, echo=F, message=F, include=F}
met_AMPO_from_DMG_max <- BXData_all %>% filter(Treatment %in% "DG") %>% 
                                         filter(Compound_name %in% c("AMPO")) %>%  
                                         filter(!uMol %in% NA) %>%  
                                         dplyr::summarize(conc_max = max(uMol))
# 0.8276532

# AMPO_from_MBOA_strong <- met_AMPO_from_MBOA_max$conc_max / 100 * 10
# 4.249294, not reached

# AMPO_from_MBOA_weak <- met_AMPO_from_MBOA_max$conc_max / 100 * 0.1
# 0.04249294, only this one relevant

```

**AAMPO formation after DIMBOA-Glc degradation**

We did not define thresholds for AAMPO formers, but recorded if AAMPO was detected in a culture.  


```{r, warning=F, echo=F, message=F}
BXData_all_phylo <- left_join(BXData_all, BXData %>% dplyr::select(Strain, Genus_Strain, Family, Phylum) %>% unique(), by = "Strain")

BXData_DG_long_5cpd <- BXData_all_phylo %>% filter(Strain != "T0") %>% 
  filter(Strain != "Standard") %>% 
  filter(Treatment %in% "DG") %>% 
  filter(Compound_name %in% c("DIMBOA-Glc", "DIM2BOA-Glc", "HMBOA-Glc", "DIMBOA", "HMBOA", "MBOA", "AMPO", "AAMPO")) %>% 
  filter(Name != "20191210_LT_LMI1_MBOA_02") %>% 
  filter(Name != "20191210_LT_LMI1_MBOA") %>% 
  filter(Name != "20191211_LT_LMI1_MBOA") %>% 
  filter(Name != "20191210_LT_NBC_DG") %>%  
  arrange(desc(Family)) %>% 
  mutate(Genus_Strain = gsub("NA_NBC", "_NBC", Genus_Strain)) %>%  
  dplyr::select(Genus_Strain, Compound_name, uMol) %>%  
  unique() %>% 
  tidyr::spread(Compound_name, uMol) %>% 
  column_to_rownames("Genus_Strain")

BXData_DG_5cpd.matrix <- as.matrix(BXData_DG_long_5cpd)
BXData_DG_5cpd.matrix[is.na(BXData_DG_5cpd.matrix)] = 0

# Change colum order
BXcols_5cpd <- c("DIMBOA-Glc", "MBOA", "AMPO", "AAMPO")

BXData_DG_5cpd.matrix <- BXData_DG_5cpd.matrix[, BXcols_5cpd]

# [, 1] DIMBOA-Glc
# [, 1] DIMBOA
# [, 1] MBOA
# [, 1] AMPO
# [, 1] AAMPO

# replace values with value with binaries
# change values - HERE STOPPPED!!!!!!!!!
## DG degraders
# strong DG degraders / no DG detected
BXData_DG_5cpd.matrix[, 1][BXData_DG_5cpd.matrix[, 1]>=0 & BXData_DG_5cpd.matrix[, 1]<=DMG_met_threshold_strong] <- 0 
# weak DG degraders / DG degraded
BXData_DG_5cpd.matrix[, 1][BXData_DG_5cpd.matrix[, 1]>=DMG_met_threshold_strong & BXData_DG_5cpd.matrix[, 1]<=DMG_met_threshold_weak] <- 4
# non-degraders of DG / DG not degraded
BXData_DG_5cpd.matrix[, 1][BXData_DG_5cpd.matrix[, 1]>=DMG_met_threshold_weak & BXData_DG_5cpd.matrix[, 1]<=700] <- 5


## MBOA formers
# non-formers / no MBOA detected
BXData_DG_5cpd.matrix[, 2][BXData_DG_5cpd.matrix[, 2]>=0 & BXData_DG_5cpd.matrix[, 2]<=MBOA_from_DMG_weak] <- 0
# weak MBOA formers / little MBOA detected
BXData_DG_5cpd.matrix[, 2][BXData_DG_5cpd.matrix[, 2]>=MBOA_from_DMG_weak & BXData_DG_5cpd.matrix[, 2]<=MBOA_from_DMG_strong] <- 1
# strong MBOA formers / lots of MBOA found
BXData_DG_5cpd.matrix[, 2][BXData_DG_5cpd.matrix[, 2]>=MBOA_from_DMG_strong & BXData_DG_5cpd.matrix[, 2]<=500] <- 2


## AMPO formers
# non-formers / no AMPO detected
BXData_DG_5cpd.matrix[, 3][BXData_DG_5cpd.matrix[, 3]>=0 & BXData_DG_5cpd.matrix[, 3]<=AMPO_from_MBOA_weak] <- 0
# weak AMPO formers / little AMPO detected
BXData_DG_5cpd.matrix[, 3][BXData_DG_5cpd.matrix[, 3]>=AMPO_from_MBOA_weak & BXData_DG_5cpd.matrix[, 3]<=AMPO_from_MBOA_strong] <- 3


## AAMPO detected
# no AAMPO detected
BXData_DG_5cpd.matrix[, 4][BXData_DG_5cpd.matrix[, 4]>=0 & BXData_DG_5cpd.matrix[, 4]<=AMPO_from_MBOA_weak] <- 0 
# AAMPO detected
BXData_DG_5cpd.matrix[, 4][BXData_DG_5cpd.matrix[, 4]>=AMPO_from_MBOA_weak ] <- 6




heatmapDIMBOAGlcFig.3_binary <- pheatmap(BXData_DG_5cpd.matrix, 
                             color = c("white", "#555CCB", "darkblue", "tomato1", "orange", "darkorange2", "firebrick1"),
                             cellwidth = 10, 
                             cutree_rows = 8, 
                             fontsize = 8, 
                             cluster_cols = FALSE, 
                             cluster_rows = FALSE, 
                             clustering_method = "average", 
                             main = "DIMBOA-Glc")


ggsave(plot = heatmapDIMBOAGlcFig.3_binary,  filename = "Fig2b.svg", width = 14, height = 21, dpi = 300, scale = 1, units = "cm")
```


## Quantitative analysis

### bargraph all compounds

```{r, warning=F, echo=F, message=F}
DG.bargraph <- BXData_all_tax %>% 
  mutate(Strain_Genus = interaction(Strain, Genus)) %>% 
  mutate(Compound_name = factor(Compound_name, levels = rev(c("DIMBOA-Glc", "DIMBOA", "DIM2BOA-Glc", "HDMBOA-Glc", "HMBOA", "HDM2BOA-Glc",  "HMBOA-Glc",   "HM2BOA-Glc", "MBOA",  "AMPO", "AAMPO",  "BOA", "APO",  "AAPO")))) %>% 
  filter(Strain != "T0") %>% 
  filter(Treatment %in% c("DG")) %>% 
  mutate(Treatment = gsub("DG", "DIMBOA-Glc", Treatment)) %>% 
  filter(Strain != "Standard") %>% 
  ggplot(aes(y = uMol, x = fct_rev(Strain_Genus))) + 
  geom_bar(aes(fill = Compound_name), position = "stack", stat = "summary", fun.y ="mean") +
  scale_fill_manual(values = level_cols_Compound) +
  theme_bw() +
  scale_y_continuous(expand = expansion(mult = c(0, 0)))+
  theme(strip.background = element_rect(color = NULL, fill= "white", linetype="solid"), strip.text.x = element_text(size = 16/.pt, margin = margin(1, 1, 1, 1)))+
  theme(axis.text.x=element_text(size = 16/.pt, angle = -90, hjust = 0, vjust = 0.5),
        axis.text.y=element_text(size = 16/.pt),
        axis.title = element_text(size = 16/.pt)) +
  coord_flip() +
  labs(x = "", y = "Concentration [µM]", fill = "") +
  geom_hline(yintercept=DMG_met_threshold_weak, color = "dimgrey")+ 
  geom_hline(yintercept=DMG_met_threshold_strong, color = "dimgrey") 

DG.bargraph

ggsave(plot = DG.bargraph, filename = "FigS5c_DG.bargraph.svg", width = 10, height = 12, dpi = 300, scale = 1, units = "cm")
```

### bargraph (A)AMPO

```{r, warning=F, echo=F, message=F}
AMPO.DG.bargraph <- BXData_all_tax %>% 
  mutate(Strain_Genus = interaction(Strain, Genus)) %>% 
  filter(Strain != "T0") %>% 
  filter(Compound_name %in% c("AMPO", "AAMPO")) %>% 
  filter(Treatment %in% c("DG")) %>% 
  mutate(Treatment = gsub("DG", "DIMBOA-Glc", Treatment)) %>% 
  filter(Strain != "Standard") %>% 
  ggplot(aes(y = uMol, x = fct_rev(Strain_Genus))) + 
  geom_bar(aes(fill = Compound_name), position = "stack", stat = "summary", fun.y ="mean") +
  facet_wrap(~Compound, ncol=2, scales="free_x") +
  scale_fill_manual(values = level_cols_Compound) +
  theme_bw() +
  scale_y_sqrt(expand=expansion(mult = c(0, 0.1))) +
  # scale_y_continuous(expand = expansion(mult = c(0, 0)))+
  theme(strip.background = element_rect(color = NULL, fill= "white", linetype="solid"), strip.text.x = element_text(size = 16/.pt, margin = margin(1, 1, 1, 1)))+
  theme(axis.text.x=element_text(size = 16/.pt, angle = -90, hjust = 0, vjust = 0.5),
        axis.text.y=element_text(size = 16/.pt),
        axis.title = element_text(size = 16/.pt)) +
  coord_flip()+
  labs(x = "", y = "Concentration [µM]", fill = "")+
  geom_hline(yintercept=AMPO_from_MBOA_weak, color = "dimgrey")

AMPO.DG.bargraph

ggsave(plot = AMPO.DG.bargraph, filename = "FigS5c_AMPO.DG.bargraph.svg", width=9, height=12, dpi = 300, scale = 1, units = "cm")
```

```{r, warning=F, echo=F, message=F}
sessionInfo()
```

