---
title: "Metabolite screen Microbacteria"
subtitle: "Thoenen et al. AMPO formation"
author: "Lisa Thönen"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_caption: yes
    toc: true
    toc_float: true
    theme: cerulean
    code_folding: show
editor_options: 
  chunk_output_type: console
---

Here we plot the results of the metabolisation assay in liquid cultures of Microbacteria from the maize root bacteria (Thoenen et al. 2023) and the AtSphere collection (Bai et al. 2015). We tested MBOA and DIMBOA-Glc. Additionally we also report growth data of all Microbacteria in MBOA-containing half-strength TSB medium to test them for MBOA tolerance and in MBOA-containing minimal medium to test them for using MBOA as sole carbon source for growth.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
rm(list = ls())
```

```{r,include=FALSE}
# install.packages("tidyverse")
# install.packages("magrittr") # for the %>% function
# install.packages("readxl") # to read excels
# install.packages("MESS") # for auc function
# install.packages("ggpubr")
# install.packages("ggpmisc")
# install.packages("pander")
# install.packages("aplot")
# install.packages("knitr")
# install.packages("ggnewscale")
# install.packages("svglite")
```

```{r, warning=F, echo=F, message=F}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

# BiocManager::install("ggtree")
```

```{r,include=FALSE}
library("tidyverse")
library("magrittr")
library("readxl")
library("MESS") 
library("ggpubr")
library("ggpmisc")
library("pander")
library("ggtree")
library("aplot")
library("knitr")
library("ggnewscale")
library("svglite")
```

```{r functions, echo=F, message=F}
give.n <- function(x){
  return(c(y = median(x)*1.05, label = length(x))) 
  # experiment with the multiplier to find the perfect position
}
```

# MBOA metabolisation 

```{r, warning=F, echo=F, message=F}
# Metadata
metadata <- read_excel("Input/210513_experimental_design_Microbacteria_MRB_At_metabolites.xlsx", sheet = 6)
# Metabolite data
met <- read_excel("Input/20211007_Microbacteria_metabolites.xlsx", sheet = 2)
# Molecular weights
mol <- read_excel("Input/Molecular_weight_BXDs.xlsx")

mol %<>% mutate(Compound = gsub("MHPA", "HMPAA", Compound)) %>% 
       mutate(Compound = gsub("MAPH", "MAP",  Compound)) 
# more info strains
database <- read_excel("Input/210513_experimental_design_Microbacteria_MRB_At_metabolites.xlsx", sheet = 2)
```

```{r, warning=F, echo=F, message=F}
met %<>% dplyr::rename(Std_conc = `Std. Conc`, ng_mL_ML = `ng/mL`) %>% 
       mutate(Compound = gsub("MHPA", "HMPAA", Compound)) %>% 
       mutate(Compound = gsub("MAPH", "MAP",  Compound))
```

```{r, warning=F, echo=F, message=F}
metadata <- left_join(metadata, database, by = c("Strain" = "Strain_ID"))
```

Quantification & calculations: using a standard curve the samples are quantified

Prepare standard curves
```{r, warning=F, echo=F, message=F}
met_Std_11 <- met %>% 
  filter(grepl('20211007_MH_LT_Bact_PMStds11x', Name)) %>% 
  filter(Type %in% "Standard") %>% 
  filter(Std_conc != "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
met_Zero <- met %>% 
  filter(grepl('20211007_MH_LT_Bact_PMStds11x', Name)) %>% 
  filter(Std_conc %in% "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
met_Std_11 <- rbind(met_Std_11, met_Zero)
met_Std_11 %<>% filter(!Std_conc %in% "NA")
```

```{r, warning=F, echo=F, message=F, fig.width=15, fig.height=12}
met_Std_11 %>% 
  filter(Compound != "HMPAA") %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x= Std_conc, y = Area)) + 
  geom_point(aes(colour = Compound), show.legend = FALSE) +
  geom_smooth(method = "lm", color="black", formula = y ~ 0 + x) +
  stat_poly_eq(formula = y ~ 0 + x,
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                parse = TRUE, label.y.npc = "top", label.x.npc = "left") +
  facet_wrap(~Compound, scales = "free") +
  labs(y = "Area",
       x = "Concentration")
```

```{r, warning=F, echo=F, message=F}
St_slopes_11 <- met_Std_11 %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>%
  group_by(Compound) %>% 
  do(lm(Area ~ 0 + Std_conc,  data = .) %>%  coef() %>%   as.data.frame() %>% slice(1))
```

```{r, warning=F, echo=F, message=F}
colnames(St_slopes_11) <- c("Compound", "Slope")
```

```{r, warning=F, echo=F, message=F}
met_Std_HMPAA <- met %>% 
  filter(grepl('20211007_MH_LT_Bact_MHPA', Name)) %>% 
  filter(Type %in% "Standard") %>% 
  filter(Std_conc != "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
met_Zero_HMPAA <- met %>% 
  filter(grepl('20211007_MH_LT_Bact_MHPA', Name)) %>% 
  filter(Std_conc %in% "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
met_Std_HMPAA <- rbind(met_Std_HMPAA, met_Zero_HMPAA)
met_Std_HMPAA %<>% filter(!Std_conc %in% "NA")
```

```{r, warning=F, echo=F, message=F}
met_Std_HMPAA %>% 
  filter(Compound %in% "HMPAA") %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x= Std_conc, y = Area)) + 
  geom_point(aes(colour = Compound), show.legend = FALSE) +
  geom_smooth(method = "lm", color="black", formula = y ~ 0 + x) +
  stat_poly_eq(formula = y ~ 0 + x,
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                parse = TRUE, label.y.npc = "top", label.x.npc = "left") +
  facet_wrap(~Compound, scales = "free") +
  labs(y = "Area",
       x = "Concentration")
```

```{r, warning=F, echo=F, message=F}
St_slopes_HMPAA <- met_Std_HMPAA %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>%
  group_by(Compound) %>% 
  do(lm(Area ~ 0 + Std_conc,  data = .) %>%  coef() %>%   as.data.frame() %>% slice(1))
```

```{r, warning=F, echo=F, message=F}
colnames(St_slopes_HMPAA) <- c("Compound", "Slope")
```

```{r, warning=F, echo=F, message=F}
St_slopes_11_HMPAA <- left_join(St_slopes_11, St_slopes_HMPAA, by = "Compound")

St_slopes_11_HMPAA_unite <- St_slopes_11_HMPAA %>% unite("slope_all", Slope.x:Slope.y, remove = FALSE)
St_slopes_11_HMPAA_unite$slope_all <- gsub("_NA", "", St_slopes_11_HMPAA_unite$slope_all)
St_slopes_11_HMPAA_unite$slope_all <- gsub("NA_", "", St_slopes_11_HMPAA_unite$slope_all)
St_slopes_11_HMPAA_unite %<>% dplyr::select(Compound, slope_all)
```

```{r, warning=F, echo=F, message=F}
met %<>% filter(Compound != "AAPO") 
```

```{r, warning=F, echo=F, message=F}
met$Name <- gsub("20211007_MH_LT_Bact_", "", met$Name)
met_t0  <- met %>% filter(Name %in% c("t0_MBOA", "t0_DMSO"))
met %<>% filter(Type %in% "Analyte")
met <- rbind(met, met_t0)
met <- left_join(met, mol, by = "Compound")
met %<>% mutate(M = as.numeric(M)) # NAs can not be mutated to numeric, a small error message pops up
```

```{r, warning=F, echo=F, message=F}
metadata %<>% dplyr::rename(Name = Sample_ID)
met <- left_join(met, metadata, by = "Name")
met %<>%
  mutate(Strain = case_when(Name == "t0_MBOA" ~ "t0",
                            Name == "t0_DMSO" ~ "t0",
                            Strain != "t0_DMSO" & Strain != "t0_MBOA" ~ Strain)) %>% 
  mutate(Treatment = case_when(Name == "t0_MBOA" ~ "MBOA",
                               Name == "t0_DMSO" ~ "DMSO",
                               Treatment != "t0_DMSO" & Treatment != "t0_MBOA" ~ Treatment))

met %<>% as.data.frame()
```

```{r, warning=F, echo=F, message=F}
met$Treatment <- factor(met$Treatment, levels = c("DMSO", "MBOA"))
```

```{r, warning=F, echo=F, message=F}
met <- left_join(met, St_slopes_11_HMPAA_unite, by = "Compound")
```

```{r, warning=F, echo=F, message=F}
met$cols <- as.character(met$Compound)
met[met$Compound == "MBOA-Glc", ]$cols <- "cornflowerblue"
met[met$Compound == "HMPMA", ]$cols <-  "darkolivegreen4"


met[met$Compound == "AMPO", ]$cols <- "darkred"
met[met$Compound == "AAMPO", ]$cols <- "firebrick1"
met[met$Compound == "HMPAA", ]$cols <- "gold2"
met[met$Compound == "MBOA", ]$cols <- "darkblue"

met[met$Compound == "HMBOA", ]$cols <- "skyblue"
met[met$Compound == "BOA", ]$cols <- "slateblue"


met[met$Compound == "DIMBOA", ]$cols <-  "tan"
met[met$Compound == "APO", ]$cols <-  "pink"
# met[met$Compound == "AAPO", ]$cols <-  "palevioletred"

temp <- data.frame(met$Compound, met$cols)
temp <- plyr::ddply(temp, .variables="met.Compound", .fun=unique)   #library(plyr)
level_cols_Compound <- as.character(temp[,2])
names(level_cols_Compound) <- temp[,1]
```

Calculate the concentrations
```{r, warning=F, echo=F, message=F}
# met$Area_sample <- as.numeric(met$Area) / 0.025 

met$Area_sample <- as.numeric(met$Area) / 0.04 
met$slope_all <- as.numeric(met$slope_all)
met$ng_sample <- met$Area_sample / met$slope_all 
met$uMol <- (met$ng_sample) / met$M
```


Quick overview of measured/detected BXs
```{r, warning=F, echo=F, message=F, fig.width=15, fig.height=12}
met %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  ggplot(aes(x = Strain, y = uMol, group = Treatment)) +
  geom_point(aes(colour = Treatment), show.legend = TRUE) +
  facet_wrap(~ Compound, scales = "free") +
  scale_colour_manual(values = c("darkblue", "darkred"))+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) 
```

```{r, warning=F, echo=F, message=F}
met %>%
  mutate(Area = as.numeric(Area)) %>%
  filter(Treatment %in% "MBOA") %>%
  ggplot(aes(x = Strain, y = uMol)) +
  geom_bar(aes(fill = Compound), stat = "identity", size = 2, show.legend = TRUE) +
  scale_fill_manual(values = level_cols_Compound)+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) +
  labs(y = "concentration [µM]" )
```

## Phylogenetic tree
Prepare tree with information on host origin

```{r, warning=F, echo=F, message=F}
### Newick tree ###
species_tree <- ape::read.tree("Input/SpeciesTree_rooted.txt")

metadata_tree <- metadata %>% mutate(Host_plant = Collection) %>% 
  mutate(Host_plant = gsub("Species_collection", "maize", Host_plant)) %>% 
  mutate(Host_plant = gsub("Maize_collection", "maize", Host_plant)) %>% 
  mutate(Host_plant = gsub("MBOA", "maize", Host_plant)) %>% 
  mutate(Host_plant = gsub("Compartment_collection", "maize", Host_plant)) %>% 
  mutate(Host_plant = gsub("AtSphere", "arabidopsis", Host_plant)) %>% 
  mutate(Host_plant = case_when(Strain == "LBN7" ~ "brassica", Strain != "LBN7" ~ Host_plant)) %>% 
  mutate(Host_plant = case_when(Strain == "LTA6" ~ "wheat", Strain != "LTA6" ~ Host_plant)) %>% 
  mutate(Host_plant = case_when(Strain == "LMS4" ~ "medicago", Strain != "LMS4" ~ Host_plant)) %>% 
  dplyr::select(Strain, AMPO_on_MBOA_plates, Host_plant) %>% 
  filter(!Strain %in% NA) %>% 
  filter(!Strain %in% "NA") %>% 
  mutate(AMPO_on_MBOA_plates = gsub("no", "negative", AMPO_on_MBOA_plates)) %>% 
  mutate(AMPO_on_MBOA_plates = gsub("yes", "positive", AMPO_on_MBOA_plates)) %>% 
  add_row(Strain = "KHB019", AMPO_on_MBOA_plates = "negative", Host_plant = "clover") %>%  #KHB019 was only characterized on plates and not in liquid culture
  mutate(Host_plant = factor(Host_plant, levels = c("maize", "arabidopsis", "brassica", "medicago", "wheat", "clover", "NA")))
  
metadata_tree_2 <- metadata_tree %>% 
        mutate(Host_plant = gsub("brassica", "other", Host_plant)) %>% 
        mutate(Host_plant = gsub("medicago", "other", Host_plant)) %>% 
        mutate(Host_plant = gsub("wheat", "other", Host_plant)) %>% 
        mutate(Host_plant = gsub("clover", "other", Host_plant))

p_host_simple <-  species_tree %>% 
      ggtree::ggtree(branch.length = "branch.length", ladderize = TRUE) %<+% metadata_tree_2 +
      geom_tippoint(aes(color=Host_plant), size = 3) + 
      geom_tiplab(align=TRUE, linetype='dashed', linesize=.3, size = 3, offset = 0.02) +
      scale_color_manual(values = c("#3e5437", "tan3", "grey"), labels = c("arabidopsis", "maize", "other"), name = "Host plant") +
      xlim(0, 2) 

# p_host_simple
```

Add AMPO phenotype on plates
```{r, warning=F, echo=F, message=F}
tax_host_pheno <- ggtree::gheatmap(p_host_simple + new_scale_fill(),
         metadata_tree %>% 
         dplyr::select(Strain, AMPO_on_MBOA_plates) %>% 
           dplyr::rename('Plate assay' = AMPO_on_MBOA_plates) %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain"),
         offset=0.3,
         width=0.2,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("lightgoldenrod", "darkred")) +
  labs(fill = "AMPO phenotype")

# tax_host_pheno
```

Add quantitative MBOA & AMPO phenotypes
```{r, warning=F, echo=F, message=F}
#Tree with quantitative MBOA&AMPO outputs
MBOA_AMPO_met_tax_host <-ggtree::gheatmap(tax_host_pheno + new_scale_fill(),
         met %>% filter(Compound %in% c("MBOA", "AMPO")) %>% 
           filter(Treatment %in% "MBOA") %>% 
           filter(!uMol %in% 0.00000000) %>%
           filter(!uMol > 600) %>%
           mutate(Compound = factor(Compound, levels = c("MBOA", "AMPO", "AAMPO", "HMPAA"))) %>% 
         dplyr::select(Strain, Compound, uMol) %>% tidyr::spread(Compound, uMol) %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain") ,
         offset=0.5,
         width=0.4,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_gradient2(low = "grey50", midpoint = 20, high = "darkred")+
  labs(#x = "Benzoxazinoid",
       #y = "Relative AUC", 
       colour = "AMPO phenotype",
       shape = "Host plant",
       fill = "Concentration [uM]") 

# MBOA_AMPO_met_tax_host
```

Prepare orthogroup data for the 5 OGs that show 100% sensitivity and specificity for AMPO phenotype on plate
```{r, warning=F, echo=F, message=F}
# add orthogroup data from DatasetS2

met_OG <- met %>% mutate(OG0001787 = 0,
               OG0002149 = 0,
               OG0002971 = 0,
               OG0002972 = 0,
               OG0002973 = 0)

met_OG %<>%
  mutate(OG0001787 = case_when(AMPO_on_MBOA_plates == "yes" ~ "2",
                            AMPO_on_MBOA_plates == "no" ~ "0")) %>% 
  mutate(OG0001787 = case_when(Strain == "LWH7" ~ "4",
                            Strain != "LWH7" ~ OG0001787)) %>%
  mutate(OG0001787 = case_when(Strain == "LWH3" ~ "3",
                            Strain != "LWH3" ~ OG0001787)) %>%
  mutate(OG0001787 = case_when(Strain == "LWS13" ~ "3",
                            Strain != "LWS13" ~ OG0001787)) %>%
  mutate(OG0002149 = case_when(AMPO_on_MBOA_plates == "yes" ~ "2",
                            AMPO_on_MBOA_plates == "no" ~ "0")) %>% 
  mutate(OG0002149 = case_when(Strain == "LWH7" ~ "3",
                            Strain != "LWH7" ~ OG0002149)) %>%
  mutate(OG0002149 = case_when(Strain == "LWH3" ~ "3",
                            Strain != "LWH3" ~ OG0002149)) %>%
  mutate(OG0002149 = case_when(Strain == "LWS13" ~ "3",
                            Strain != "LWS13" ~ OG0002149)) %>%
  mutate(OG0002149 = case_when(Strain == "LWH13" ~ "1",
                            Strain != "LWH13" ~ OG0002149)) %>%
  mutate(OG0002149 = case_when(Strain == "LM3X" ~ "1",
                            Strain != "LM3X" ~ OG0002149)) %>%
  mutate(OG0002149 = case_when(Strain == "LMB2" ~ "1",
                            Strain != "LMB2" ~ OG0002149)) %>%
  mutate(OG0002149 = case_when(Strain == "LMX7" ~ "1",
                            Strain != "LMX7" ~ OG0002149)) %>%
  mutate(OG0002149 = case_when(Strain == "LWO14" ~ "1",
                            Strain != "LWO14" ~ OG0002149)) %>%
  mutate(OG0002971 = case_when(AMPO_on_MBOA_plates == "yes" ~ "1",
                           AMPO_on_MBOA_plates == "no" ~ "0")) %>% 
  mutate(OG0002972 = case_when(AMPO_on_MBOA_plates == "yes" ~ "1",
                            AMPO_on_MBOA_plates == "no" ~ "0")) %>% 
  mutate(OG0002973 = case_when(AMPO_on_MBOA_plates == "yes" ~ "1",
                            AMPO_on_MBOA_plates == "no" ~ "0"))  
  
```

```{r, warning=F, echo=F, message=F}
met_OG_selected <- met_OG %>% 
         dplyr::select(Strain, OG0001787, OG0002149, OG0002971, OG0002972, OG0002973) %>% 
           mutate(OG0001787 = replace_na(OG0001787, "0")) %>% 
           mutate(OG0002149 = replace_na(OG0002149, "0")) %>% 
           mutate(OG0002971 = replace_na(OG0002971, "0")) %>% 
           mutate(OG0002972 = replace_na(OG0002972, "0")) %>% 
           mutate(OG0002973 = replace_na(OG0002973, "0")) %>% 
           add_row(Strain = "KHB019", OG0001787 = "0", OG0002149 = "0", OG0002971 = "0", OG0002972 = "0", OG0002973 = "0") %>% unique() #KHB019 is added manually because this strain was not tested in the metabolomics experiment and hence was not present in the dataframe used to prepare this one.
```


Plot AMPO phenotype in plate assay along with quantitative MBOA and AMPO concentrations measured for liquid cultures grown in half-strength TSB supplemented with 500 uM MBOA. Orthogroups showing 100% sensitivity and specificity are shown on the side. Note that KHB019 was not assessed in liquid culture and therefore is shaded in grey for MBOA and AMPO.
```{r, warning=F, echo=F, message=F}
MBOA_AMPO_met_tax_shape_OG_host <- ggtree::gheatmap(MBOA_AMPO_met_tax_host + new_scale_fill(),
         met_OG_selected %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain"),
         offset=0.8,
         width=1,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("gray100", "gray80", "gray60", "gray40", "gray20", "white")) +
  labs(fill = "Orthogroup copies")

MBOA_AMPO_met_tax_shape_OG_host

```


```{r, warning=F, echo=F, message=F}
met_bar_phylo <- met %>% filter(Treatment %in% "MBOA") %>% filter(!Compound %in% c("AMPO", "AAMPO")) %>% dplyr::select(Strain, Compound, uMol) %>% pivot_wider(names_from = Compound, values_from = uMol) 

#for exporting of the NBC condition, place it as KHB019 which was not measured. Then correct in Illustrator by moving the columns for KHB019 to NBC and removing from KHB019.
met_bar_phylo$Strain <- gsub("NBC", "KHB019", met_bar_phylo$Strain)
met_bar_phylo %<>% as.data.frame()

met_phylo <- met %>% filter(Treatment %in% "MBOA") %>% filter(Strain != "t0") %>% dplyr::select(Strain, Compound, uMol)
met_phylo$Strain <- gsub("NBC", "KHB019", met_phylo$Strain)
```

Figure S8a: BX compounds measured after incubation in half-strength TSB supplemented with 500 uM MBOA.
```{r, warning=F, echo=F, message=F}
p_met <- met_phylo %>% 
         mutate(uMol = case_when(Strain == "KHB019" ~ 0,
                                  Strain != "KHB019" ~ uMol)) %>% 
        ggplot(aes(y = Strain, x = uMol)) +
        geom_bar(aes(fill = Compound), stat = "identity") +
        # scale_fill_jco()+
        scale_fill_manual(values = level_cols_Compound)+
        theme_tree2() +
        labs(x = "Concentration [uM]")

plot_tax_met_micro <- p_met %>% 
  insert_left(p_host_simple, width = 0.5)


plot_tax_met_micro

ggsave(plot = plot_tax_met_micro, filename = "FigS8a_Micro_MBOA_tree.svg", width = 15, height = 15, dpi = 300, scale = 1, units = "cm")

# To export the no bacteria control NBC use it at position for KHB019, which was not included in the assays and then add it in Illustrator
p_met_NBC_asKHB019 <- met_phylo %>% 
        ggplot(aes(y = Strain, x = uMol)) +
        geom_bar(aes(fill = Compound), stat = "identity") +
        scale_fill_manual(values = level_cols_Compound)+
        theme_tree2() +
        labs(x = "Concentration [uM]")
p_met_NBC_asKHB019_tax <- p_met_NBC_asKHB019 %>% 
  insert_left(p_host_simple, width = 0.5)

# ggsave(plot = p_met_NBC_asKHB019_tax, filename = "FigS8a_Micro_MBOA_tree_NBC_as_KHB019.svg", width = 15, height = 15, dpi = 300, scale = 1, units = "cm")
```

# DIMBOA-Glc metabolisation

```{r, warning=F, echo=F, message=F}
metadataDA <- read.csv("Input/Microbacteria_DMSO_DMG_sample_list.csv")

metadataDA <- left_join(metadataDA, database, by = c("Strain" = "Strain_ID"))

metDA <- read.csv("Input/Microbact_DMG_metabolites_quantified.csv")


metDA %<>% dplyr::rename(Std_conc = Std..Conc, ng_mL_ML = ng.mL) %>% 
       mutate(Compound = gsub("MHPA", "HMPAA", Compound)) %>% 
       mutate(Compound = gsub("MAPH", "MAP",  Compound))
```

Quantification & calculations: using a standard curve the samples are quantified

Prepare standard curves
```{r, warning=F, echo=F, message=F}
metDA_Std_11 <- metDA %>% 
  filter(grepl('20220203_MH_LT_Microbact_metabolites_PM11xStds', Name)) %>% 
  filter(Type %in% "Standard") %>% 
  filter(Std_conc != "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
metDA_Zero <- metDA %>% 
  filter(grepl('20220203_MH_LT_Microbact_metabolites_PM11xStds', Name)) %>% 
  filter(Std_conc %in% "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
metDA_Std_11 <- rbind(metDA_Std_11, metDA_Zero)
metDA_Std_11 %<>% filter(!Std_conc %in% "NA")
```

```{r, warning=F, echo=F, message=F, fig.width=15, fig.height=12}
metDA_Std_11 %>% 
  filter(!Compound %in% c("AAPO", "HMPAA", "DIBOA-Glc", "HBOA-Glc", "HM2BOA-Glc", "HMBOA-Glc", "MAPH", "MAP")) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x= Std_conc, y = Area)) + 
  geom_point(aes(colour = Compound), show.legend = FALSE) +
  geom_smooth(method = "lm", color="black", formula = y ~ 0 + x) +
  stat_poly_eq(formula = y ~ 0 + x,
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                parse = TRUE, label.y.npc = "top", label.x.npc = "left") +
  facet_wrap(~Compound, scales = "free") +
  labs(y = "Area",
       x = "Concentration")
```

 
```{r, warning=F, echo=F, message=F}
St_slopes_11_DA <- metDA_Std_11 %>% 
  filter(!Compound %in% c("AAPO", "HMPAA", "DIBOA-Glc", "HBOA-Glc", "HM2BOA-Glc", "HMBOA-Glc", "MAP")) %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>%
  group_by(Compound) %>% 
  do(lm(Area ~ 0 + Std_conc,  data = .) %>%  coef() %>%   as.data.frame() %>% slice(1))
```

```{r, warning=F, echo=F, message=F}
colnames(St_slopes_11_DA) <- c("Compound", "Slope")
```

```{r, warning=F, echo=F, message=F}
metDA_Std_HMPAA <- metDA %>% 
  filter(grepl('20220203_MH_LT_Microbact_metabolites_MHPA', Name)) %>% 
  filter(Type %in% "Standard") %>% 
  filter(Std_conc != "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
metDA_Zero_HMPAA <- metDA %>% 
  filter(grepl('20220203_MH_LT_Microbact_metabolites_MHPA', Name)) %>% 
  filter(Std_conc %in% "0.000") %>% 
  dplyr::select(Std_conc, Area, Compound)
```

```{r, warning=F, echo=F, message=F}
metDA_Std_HMPAA <- rbind(metDA_Std_HMPAA, metDA_Zero_HMPAA)
metDA_Std_HMPAA %<>% filter(!Std_conc %in% "NA")
```

```{r, warning=F, echo=F, message=F}
metDA_Std_HMPAA %>% 
  filter(Compound %in% "HMPAA") %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x= Std_conc, y = Area)) + 
  geom_point(aes(colour = Compound), show.legend = FALSE) +
  geom_smooth(method = "lm", color="black", formula = y ~ 0 + x) +
  stat_poly_eq(formula = y ~ 0 + x,
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                parse = TRUE, label.y.npc = "top", label.x.npc = "left") +
  facet_wrap(~Compound, scales = "free") +
  labs(y = "Area",
       x = "Concentration")
```

```{r, warning=F, echo=F, message=F}
St_slopes_HMPAA_DA <- metDA_Std_HMPAA %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  mutate(Std_conc = as.numeric(Std_conc)) %>%
  group_by(Compound) %>% 
  do(lm(Area ~ 0 + Std_conc,  data = .) %>%  coef() %>%   as.data.frame() %>% slice(1))
```

```{r, warning=F, echo=F, message=F}
colnames(St_slopes_HMPAA_DA) <- c("Compound", "Slope")
```
join 11 stds and HMPAA
```{r, warning=F, echo=F, message=F}
St_slopes_11_HMPAA_DA_unite <- rbind(St_slopes_11_DA, St_slopes_HMPAA_DA %>% filter(Compound %in% "HMPAA"))
```

remove the compounds which are not reliable quantified
```{r, warning=F, echo=F, message=F}
metDA %<>% filter(Compound != "AAPO") 
```

```{r, warning=F, echo=F, message=F}
metDA$Name <- gsub("20220203_MH_LT_Microbact_metabolites_", "", metDA$Name)
metDA %<>% filter(Type %in% "Analyte")
metDA <- left_join(metDA, mol, by = "Compound")
metDA %<>% mutate(M = as.numeric(M)) # NAs can not be mutated to numeric, a small error message pops up
```

```{r, warning=F, echo=F, message=F}
metadataDA %<>% dplyr::rename(Name = Sample)

metDA <- left_join(metDA, metadataDA, by = "Name")
```



```{r, warning=F, echo=F, message=F}
metDA$Treatment <- factor(metDA$Treatment, levels = c("D", "G", "A")) # D corresponds to DMSO and G to DIMBOA-Glc

```

```{r, warning=F, echo=F, message=F}
metDA <- left_join(metDA, St_slopes_11_HMPAA_DA_unite, by = "Compound")
```

color code Compounds
```{r, warning=F, echo=F, message=F}
metDA$cols <- as.character(metDA$Compound)
metDA[metDA$Compound == "AMPO", ]$cols <- "darkred"
metDA[metDA$Compound == "AAMPO", ]$cols <- "firebrick1"
metDA[metDA$Compound == "HMPAA", ]$cols <- "gold2"
metDA[metDA$Compound == "MBOA", ]$cols <- "darkblue"
metDA[metDA$Compound == "MAP", ]$cols <- "khaki"

metDA[metDA$Compound == "HMBOA", ]$cols <- "skyblue"
metDA[metDA$Compound == "BOA", ]$cols <- "slateblue"
metDA[metDA$Compound == "MBOA-Glc", ]$cols <- "darkmagenta"


metDA[metDA$Compound == "HMPMA", ]$cols <-  "slategray"
metDA[metDA$Compound == "DIMBOA", ]$cols <-  "tan"
metDA[metDA$Compound == "APO", ]$cols <-  "pink"
# metDA[metDA$Compound == "AAPO", ]$cols <-  "pink3"

metDA[metDA$Compound == "DIMBOA-Glc", ]$cols <-  "darkorange2"
metDA[metDA$Compound == "HDMBOA-Glc", ]$cols <-  "thistle3"
metDA[metDA$Compound == "HBOA-Glc", ]$cols <-  "slategray"
metDA[metDA$Compound == "DIBOA-Glc", ]$cols <-  "slategray"
metDA[metDA$Compound == "HMBOA-Glc", ]$cols <-  "skyblue2"
metDA[metDA$Compound == "HM2BOA-Glc", ]$cols <-  "skyblue4"
metDA[metDA$Compound == "DIM2BOA-Glc", ]$cols <-  "sandybrown"
metDA[metDA$Compound == "HDM2BOA-Glc", ]$cols <-  "thistle1"

temp <- data.frame(metDA$Compound, metDA$cols)
temp <- plyr::ddply(temp, .variables="metDA.Compound", .fun=unique)   #library(plyr)
level_cols_Compound <- as.character(temp[,2])
names(level_cols_Compound) <- temp[,1]
```


Dilutions: 
150 + 350 = 500 
50 + 700 = 750 
500 / 150 = 3.33333
750 / 50 = 15

1/((500/150)*15) = 0.02

for some reason the MBOA concentrations are a bit too high (factor 1.273141 controlled with t0 MBOA sample) --> ev because of precipitaion issue?

```{r, warning=F, echo=F, message=F}
metDA$Area_sample <- as.numeric(metDA$Area) / 0.025 
metDA$ng_sample <- metDA$Area_sample / metDA$Slope 
metDA$uMol <- (metDA$ng_sample) / metDA$M
```

Explore raw data

```{r, warning=F, echo=F, message=F, include = F}
metDA %>% 
  filter(!is.na(uMol)) %>% 
  mutate(Compound = as.factor(Compound)) %>% 
  ggplot(aes(x = Strain, y = uMol, group = Treatment)) +
  geom_point(aes(colour = Treatment), show.legend = TRUE) +
  facet_wrap(~ Compound, scales = "free") +
  scale_colour_manual(values = c("darkblue", "darkorange2", "maroon"))+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) +
  labs(y = "uM")

```

```{r, warning=F, echo=F, message=F}
metDA %>% 
  filter(!is.na(uMol)) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  ggplot(aes(x = Strain, y = uMol, group = Treatment)) +
  geom_bar(aes(fill = Treatment), stat = "identity", size = 2, show.legend = TRUE) +
  facet_wrap(~ Compound, scales = "fixed") +
  scale_fill_manual(values = c("darkblue", "darkorange2", "maroon"))+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) +
  labs(y = "uM")
```


```{r, warning=F, echo=F, message=F}
metDA %>% 
  filter(!Compound %in% c("DIM2BOA-Glc", "HDM2BOA-Glc")) %>% 
  mutate(Area = as.numeric(Area)) %>% 
  filter(Treatment %in% "G") %>% 
  ggplot(aes(x = Strain, y = uMol)) +
  geom_bar(aes(fill = Compound), stat = "identity", size = 2, show.legend = TRUE) +
  scale_fill_manual(values = level_cols_Compound)+
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) +
  labs(title = "DIMBOA-Glc")



```

## Phylogenetic tree
```{r, warning=F, echo=F, message=F}
### Newick tree ###
species_tree <- ape::read.tree("Input/SpeciesTree_rooted.txt")

metadata_tree <- metadata %>% dplyr::select(Strain, AMPO_on_MBOA_plates) %>% filter(!Strain %in% NA)

p <-  species_tree %>% 
      ggtree::ggtree(branch.length = "none", ladderize = TRUE) %<+% metadata_tree + 
      geom_tippoint(aes(color=AMPO_on_MBOA_plates)) + 
      geom_tiplab(size=3) + 
      scale_color_manual(values = c("gold", "darkred"))
```


```{r, warning=F, echo=F, message=F}
metDA_bar_phylo <- metDA %>% filter(Treatment %in% "G") %>% dplyr::select(Strain, Compound, uMol) %>% pivot_wider(names_from = Compound, values_from = uMol) 

#for exporting of the NBC condition, place it as KHB019 which was not measured. Then correct in Illustrator by moving the columns for KHB019 to NBC and removing from KHB019.
metDA_bar_phylo$Strain <- gsub("NBC", "KHB019", metDA_bar_phylo$Strain)
metDA_bar_phylo %<>% as.data.frame()

metDA_phylo <- metDA %>% filter(Treatment %in% "G") %>% filter(Strain != "T0") %>% filter(!Compound %in% c("HDM2BOA-Glc", "DIM2BOA-Glc", "MAP")) %>% dplyr::select(Strain, Compound, uMol)
metDA_phylo$Strain <- gsub("NBC", "KHB019", metDA_phylo$Strain) # for exporting of NBC condition because KHB019 was not measured
```

Figure S8b: BX compounds measured after incubation in half-strength TSB supplemented with DIMBOA-Glc
```{r, warning=F, echo=F, message=F}
# to get the NBC, use it as KHB019
p_metDA_DMG_NBC_asKHB019 <- metDA_phylo %>% 
        mutate(Compound_name = factor(Compound)) %>% 
        mutate(Compound_name = fct_relevel(Compound_name, "DIMBOA-Glc", after = Inf)) %>% # this makes sure that DIMBOA-Glc comes first.  
        ggplot(aes(y = Strain, x = uMol)) +
        geom_bar(aes(fill = Compound_name), stat = "identity") +
        xlim(0, 800)+
        scale_fill_manual(values = level_cols_Compound, drop = F)+
        theme_tree2() +
        labs(x = "Concentration [uM]")

p_metDA_DMG_NBC_asKHB019_micro <- p_metDA_DMG_NBC_asKHB019 %>% 
  insert_left(p_host_simple, width = 0.5)  

# ggsave(plot = plot_tax_metDA_DMG_micro, filename = "FigS8b_Micro_DIMBOA_Glc_tree_NBC_as_KHB019.svg",width = 15, height = 15, dpi = 300, scale = 1, units = "cm")


# to get the NBC, use it as KHB019
p_metDA_DMG <- metDA_phylo %>% 
        mutate(Compound_name = factor(Compound)) %>% 
        mutate(Compound_name = fct_relevel(Compound_name, "DIMBOA-Glc", after = Inf)) %>% # this makes sure that DIMBOA-Glc comes first.  
        mutate(uMol = ifelse(Strain == "KHB019", 0, uMol)) %>% 
        ggplot(aes(y = Strain, x = uMol)) +
        geom_bar(aes(fill = Compound_name), stat = "identity") +
        xlim(0, 800)+
        scale_fill_manual(values = level_cols_Compound, drop = F)+
        theme_tree2() +
        labs(x = "Concentration [uM]")

plot_tax_metDA_DMG_micro <- p_metDA_DMG %>% 
  insert_left(p_host_simple, width = 0.5)  

plot_tax_metDA_DMG_micro

ggsave(plot = plot_tax_metDA_DMG_micro, filename = "FigS8b_Micro_DIMBOA_Glc_tree.svg",width = 15, height = 15, dpi = 300, scale = 1, units = "cm")


```


# Metabolite categories for comparative genomics

```{r, warning=F, echo=F, message=F}
met_all <- rbind(met %>% dplyr::select(-Nr, -Nr.x, -T, -n, -Nr.y) %>% dplyr::rename(Slope = slope_all, IS.Area = 'IS Area', Detection.Flags = 'Detection Flags'), metDA %>% dplyr::select(-X.1, -X.x, -X., -random, -Nr.x, -Nr.y, -X0, -Treat, -X.y) %>% dplyr::rename('%Dev' = X.Dev))
```

```{r, warning=F, echo=F, message=F}
met_all_quant <- met_all %>% dplyr::select(Strain, Treatment, Compound, uMol) %>% filter(Treatment != "DMSO") %>% mutate(Treatment_Compound = interaction(Treatment, Compound, sep = "_")) %>% dplyr::select(-Treatment, -Compound) %>% filter(!Strain %in% c("t0", "T0")) %>% filter(Treatment_Compound != "MBOA_AAMPO") %>% unique() %>% replace(is.na(.), 0) %>%  pivot_wider(names_from = Treatment_Compound, values_from = uMol, values_fn = mean)
```

**MBOA**
```{r, warning=F, echo=F, message=F}
## MBOA metabolization < 400
max_MBOA <- met_all_quant$MBOA_MBOA %>% max()
# [1] 516.2433

## AMPO formation > 2.5
max_AMPO <-met_all_quant$MBOA_AMPO %>% max()
# [1] 302.6399

## HMPAA formation > 25
max_HMPAA <-met_all_quant$MBOA_HMPAA %>% max()
# [1] 44.06674

tibble(max_MBOA, max_AMPO, max_HMPAA) %>% knitr::kable(caption = "MBOA metabolisation")
```

**DIMBOA-Glc metabolization**
```{r, warning=F, echo=F, message=F}
## DIMBOA-Glc metabolization < 400 (LMI1x 322 uM)
max_DIMBOA_Glc <- met_all_quant$`G_DIMBOA-Glc` %>% max()
# [1] 575.2392

## MBOA formation > 40 (LMI1x 43.7 uM)
max_MBOA <- met_all_quant$G_MBOA %>% max()
# [1] 120.9501

## AMPO formation DMG > 2.5
max_AMPO <- met_all_quant$G_AMPO %>% max()
# [1] 7.355668

tibble(max_DIMBOA_Glc, max_MBOA, max_AMPO) %>% knitr::kable(caption = "DIMBOA-Glc metabolisation")
```

**Groups**
```{r, warning=F, echo=F, message=F}
met_all_quali <- met_all_quant %>%
  mutate(MBOA_met = case_when(MBOA_MBOA > 400 ~ FALSE,
                              MBOA_MBOA < 400 ~ TRUE)) %>%
  mutate(MBOA_AMPO = case_when(MBOA_AMPO > 2.5 ~ TRUE,
                               MBOA_AMPO < 2.5 ~ FALSE)) %>%
  mutate(MBOA_HMPAA = case_when(MBOA_HMPAA > 15 ~ TRUE,
                              MBOA_HMPAA < 15 ~ FALSE)) %>%
  mutate(DIMBOA_Glc_met = case_when(`G_DIMBOA-Glc` > 400 ~ FALSE,
                              `G_DIMBOA-Glc` < 400 ~ TRUE)) %>%
  mutate(DMG_MBOA = case_when(G_MBOA > 40 ~ TRUE,
                              G_MBOA < 40 ~ FALSE)) %>%
  mutate(DMG_AMPO = case_when(G_AMPO > 2.5 ~ TRUE,
                              G_AMPO < 2.5 ~ FALSE)) %>%
  dplyr::select(Strain, MBOA_met, MBOA_AMPO, MBOA_HMPAA, DIMBOA_Glc_met, DMG_MBOA, DMG_AMPO) %>%    left_join(., database %>% dplyr::select(Strain_ID, AMPO_on_MBOA_plates), by = c("Strain" = "Strain_ID"))
```

# Growth Microbacteria TSB (Metabolite samples) & MM (Carbon source)

```{r, include=FALSE}
Strain_Data <- read_excel("Input/210513_experimental_design_Microbacteria_MRB_At_metabolites.xlsx", sheet = 5)
Strain_Data 
database <- read_excel("Input/210513_experimental_design_Microbacteria_MRB_At_metabolites.xlsx", sheet = 2)
Strain_Data <- left_join(Strain_Data, database, by = c("Strain" = "Strain_ID")) %>% dplyr::select(-'old name', -'Possible_species/best_hit') 
Strain_Data$Well_Plate <- interaction(Strain_Data$Well, Strain_Data$Plate, sep = "_")

Strain_Data %<>% arrange(., desc(Strain))
Strain_Data$Block <- rep(c(1:3), times = 64)
```

```{r, warning=FALSE, include=FALSE}
results1 <- data_frame()

# Provide a lookup table for plate information, treatments are spread accross sheets in the table to be read.
# in the first column describe the plates and in the second column the experiment (rep(times = # plates - 2))
treatlist <- data_frame("Treat" =
              c("Dummy",
                "Glc_30000_MM_B_2",
                "Glc_30000_MM_A_2",
                "Glc_30000_MM_B_1",
                "Glc_30000_MM_A_1",
                "MBOA_MM_B_2",
                "MBOA_MM_A_2",
                "MBOA_MM_B_1",
                "MBOA_MM_A_1",
                "DMSO_MM_B_2",
                "DMSO_MM_A_2",
                "DMSO_MM_B_1",
                "DMSO_MM_A_1",
                "MBOA_TSB_B_2",
                "MBOA_TSB_A_2",
                "MBOA_TSB_B_1",
                "MBOA_TSB_A_1",
                "DMSO_TSB_B_2",
                "DMSO_TSB_A_2",
                "DMSO_TSB_B_1",
                "DMSO_TSB_A_1",
                "no_TSB_MM_no_1",
                "Dummy"),
 "Replicate" = c("NULL", rep(c("2", "2","1", "1"), times = 5), "1", "NULL"), # times = all plates but without the dummy plates
 "Plate" = c("NULL", rep(c("B", "A"), times = 10), "NULL", "NULL"), 
# times = all plates but without the dummy plates
 "Media" = c("NULL", rep("MM", times = 12), rep ("TSB", times =8), "TSB_MM", "NULL"),
"Chem" = c("NULL", rep("Glucose", times = 4), rep("MBOA", times = 4), rep("DMSO", times = 4), rep("MBOA", times = 4), rep("DMSO", times = 4), "no", "NULL"), 
"Run" = c("NULL", rep("R1", times = 21), "NULL"))


#Initialize / clear df

#Loop through sheets.

for (i in 1:nrow(treatlist)) {
  tmp <- read_excel("Input/Microbacteria_metabolite_210826_repe_data_reader.xlsx", sheet = i, range = "B26:CU70") %>%   # change for duration runs
    dplyr::select(-`T° 600`) %>% # Drop Temperature
    #pivot_longer(names_to = "Well", values_to = "Density") %>%  #Make long
    gather(key = "Well", value = "Density", 2:ncol(.)) %>%
    mutate(Treat = as.character(treatlist[i,1]), #Add treatment
    Replicate = as.character(treatlist[i,2]),
    Plate = as.character(treatlist[i,3]),
    Media = as.character(treatlist[i,4]),
    Chem = as.character(treatlist[i,5]),
    Run = as.character(treatlist[i,6]),
    Well_Plate = interaction(Well, Plate, sep = "_")) #Add Replicate
results1 <- bind_rows(results1, tmp)
}
```

```{r, warning=FALSE, include=FALSE}
results2 <- data_frame()

# Provide a lookup table for plate information, treatments are spread accross sheets in the table to be read.
# in the first column describe the plates and in the second column the experiment (rep(times = # plates - 2))
treatlist <- data_frame("Treat" =
              c("Dummy",
                "Glc_30000_MM_B_2",
                "Glc_30000_MM_A_2",
                "Glc_30000_MM_B_1",
                "Glc_30000_MM_A_1",
                "MBOA_MM_B_2",
                "MBOA_MM_A_2",
                "MBOA_MM_B_1",
                "MBOA_MM_A_1",
                "DMSO_MM_B_2",
                "DMSO_MM_A_2",
                "DMSO_MM_B_1",
                "DMSO_MM_A_1",
                "MBOA_TSB_B_2",
                "MBOA_TSB_A_2",
                "MBOA_TSB_B_1",
                "MBOA_TSB_A_1",
                "DMSO_TSB_B_2",
                "DMSO_TSB_A_2",
                "DMSO_TSB_B_1",
                "DMSO_TSB_A_1",
                "no_TSB_MM_no_1",
                "Dummy"),
 "Replicate" = c("NULL", rep(c("2", "2","1", "1"), times = 5), "1", "NULL"), # times = all plates but without the dummy plates
 "Plate" = c("NULL", rep(c("B", "A"), times = 10), "NULL", "NULL"), 
# times = all plates but without the dummy plates
 "Media" = c("NULL", rep("MM", times = 12), rep ("TSB", times =8), "TSB_MM", "NULL"),
"Chem" = c("NULL", rep("Glucose", times = 4), rep("MBOA", times = 4), rep("DMSO", times = 4), rep("MBOA", times = 4), rep("DMSO", times = 4), "no", "NULL"), 
"Run" = c("NULL", rep("R2", times = 21), "NULL"))


#Initialize / clear df

#Loop through sheets.

for (i in 1:nrow(treatlist)) {
  tmp <- read_excel("Input/Microbacteria_metabolite_210930_Run2.xlsx", sheet = i, range = "B26:CU70") %>%   # change for duration runs
    dplyr::select(-`T° 600`) %>% # Drop Temperature
    #pivot_longer(names_to = "Well", values_to = "Density") %>%  #Make long
    gather(key = "Well", value = "Density", 2:ncol(.)) %>%
    mutate(Treat = as.character(treatlist[i,1]), #Add treatment
    Replicate = as.character(treatlist[i,2]),
    Plate = as.character(treatlist[i,3]),
    Media = as.character(treatlist[i,4]),
    Chem = as.character(treatlist[i,5]),
    Run = as.character(treatlist[i,6]),
    Well_Plate = interaction(Well, Plate, sep = "_")) #Add Replicate
results2 <- bind_rows(results2, tmp)
}
```

```{r, warning=FALSE, include=FALSE}
results <- rbind(results1, results2)
```

```{r, warning=FALSE, include=FALSE}
###
###Get times based on difftime. 
results %<>%
  filter(Treat != "Dummy" ) %>%
  filter(Treat != "Dummy") %>%
  filter(Density != "NA") %>%
  group_by(Replicate, Well_Plate, Treat) %>%
  mutate(mins = difftime(Time, Time[1], units = "mins") %>% as.numeric(),
         hrs = difftime(Time, Time[1], units = "hours") %>% as.numeric())  %>% 
  ###Merge with metadata. Well Positions are Part of the stocklist (clean), other metadata comes from the ZmSphere.
  left_join(., Strain_Data, by = "Well_Plate") # 

results %<>% dplyr::select(-Well.y, -Plate.x, -Time, -Row, -Column) %>% dplyr::rename(Well = Well.x, Plate = Plate.y) %>%  as.data.frame() 

results %<>% mutate(Row = Well %>% str_extract(.,"[A-H]"),
                    Column = Well %>% str_extract(.,"[0-9]+") %>% as.numeric()) 

```

```{r warning=FALSE, include=FALSE}
###Make density increase relative to first timepoint
results %<>%
  filter(!Strain %in% c("Ignore","None")) %>%
  group_by(Replicate, Well_Plate, Treat) %>%
  mutate(Density_increase = Density - Density[1]) %>%
  ungroup %>% as.data.frame()
```

```{r, warning=FALSE, include=FALSE}
results_raw <- results
```

## Growth curves of all plates single treatments
```{r, warning=FALSE, message=F, echo=F, include = FALSE}
results %>%  filter(Treat %in% "no_TSB_MM_no_1") %>% 
  ggplot(aes(x=hrs, y= Density_increase)) +
  geom_line(aes(color = Run), size = 1) +
  facet_grid(Row ~ Column, scales = "fixed") +
  theme_bw() +
  labs(x = "Time [hours]",
       y = "OD 600", 
       title = "Empty plate",
       subtitle = "control for contamination", 
       color = "Run")
```

```{r, warning=FALSE, message=F, echo=F, include = FALSE}
results %>%  
  filter(Media %in% "MM") %>% 
  filter(!Strain %in% "no") %>% 
  filter(!Strain %in% "NBC") %>% 
  ggplot(aes(x=hrs, y= Density_increase)) +
  geom_line(aes(linetype = interaction(Run, Rep, Replicate), color = Chem), size = 1) +
  facet_wrap(~ Strain, scales = "fixed") +
  theme_bw() +
  labs(x = "Time [hours]",
       y = "OD 600", 
       title = "Growth curves of all strains in MM supplemented with carbon source",
       color = "Run")
```

```{r, warning=FALSE, message=F, echo=F, include = FALSE}
results %>%  filter(Media %in% "MM") %>% 
  filter(Chem %in% c("DMSO", "MBOA")) %>% 
  ggplot(aes(x=hrs, y= Density_increase, color = AMPO_on_MBOA_plates)) +
  geom_line(aes(linetype = interaction(Run, Rep, Replicate), color = Chem), size = 1, show.legend = FALSE) +
  facet_wrap( ~ Strain, scales = "fixed") +
  theme_bw() +
  labs(x = "Time [hours]",
       y = "OD 600", 
       title = "Variation: Glc_30000_MM_A_1 & Glc_30000_MM_A_2",
       subtitle = "growth in mix - Plate A + Plate B", 
       color = "Run")
```

## AUC 

To quantify the total bacterial growth over time, we calculate the total area under the curve. This is done with the function "auc()". Before that, negative values for density increase were filtered out (i.e., strains or replicates that didn't grow). For comparison between treatments, the total AUC (calculated using density increase) and AUC_raw (calculated using raw density) is normalized with the AUC of the strain grown in the control treatment (no chemicals added, just normal growth media with DMSO). 

```{r, warning=FALSE, include=FALSE, echo=F, include=F}
results_AUC <- results %>% 
  filter(Treat != "Dummy") %>% 
  filter(Density_increase > 0) %>% # removes 'negative' growth curves
  filter(!Strain %in% c("no", NA)) %>% 
  group_by(Strain, Well, Chem, Media, Plate, Replicate, Run) %>%
  dplyr::summarize(AUC = auc(hrs, Density_increase), AUC_raw = auc(hrs, Density)) %>%
  ungroup %>% 
  mutate(Chem = as.factor(Chem),
         Plate = as.factor(Plate),
         Replicate = as.factor(Replicate),
         Media = as.factor(Media),
         Well = as.factor(Well),
         Run = as.factor(Run)) %>% 
  mutate(AUC_pos = abs(AUC)) %>% 
  mutate(Strain_Well_Media_Chem_Plate_Replicate_Run = paste(Strain, Well, Media, Chem, Plate, Replicate, Run, sep = "_")) %>%
  #Split by Replicate
  base::split(., .$Strain_Well_Media_Chem_Plate_Replicate_Run) %>% #Make each strain / rep combination its own data.frane
  map_dfr(~
  mutate(.x, AUC_norm = AUC_pos / AUC_pos[length(AUC_pos)])) %>%
  #in each df, divide the AUC by the last AUC value; since Control does not start with a number, it is the last treatment
  left_join(. , Strain_Data %>% 
  dplyr::select(Strain, Phylum, Family, Genus, AMPO_on_MBOA_plates, Collection) %>% unique(), by = "Strain") %>% 
  unique() %>% 
  as.data.frame() #transform into df


## Save table with AUC
#write_rds(results_AUC, paste0("results_AUC.rds"))
```

```{r, echo=F, warning=F, message=F, fig.width=20, fig.height=15, include=F}
results_AUC %>%  
  ggplot(aes(x= interaction(as.factor(Chem), Media), y= AUC)) +
  geom_boxplot(aes(fill = AMPO_on_MBOA_plates), position=position_dodge()) + 
  geom_jitter(aes(shape = Run), width = 0.1) + 
  ggpubr::stat_compare_means(label = "p.signif", method = "t.test", ref.group = "DMSO.MM", label.y.npc = 0.8) +
  ylim(0, 60) +
  facet_wrap(~Strain, scales = "fixed") +
  scale_fill_manual(values = c("cornflowerblue", "darkred")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0.5)) +
    labs(y = "AUC", 
         x = "",
         title = "Growth of strains in minimal media",
         colour = "AMPO on plate")
```

### TSB

```{r, echo=F, warning=F, message=F, fig.width=20, fig.height=15, include=F}
results_AUC %>%  
  filter(Media %in% "TSB") %>% 
  ggplot(aes(x= interaction(as.factor(Chem)), y= AUC)) +
  geom_boxplot(aes(fill = AMPO_on_MBOA_plates), position=position_dodge()) + 
  geom_jitter(width = 0.1) + 
  ggpubr::stat_compare_means(label = "p.signif", method = "t.test", ref.group = "DMSO", label.y = 60) +
  facet_wrap(~Strain, scales = "fixed") +
  scale_fill_manual(values = c("cornflowerblue", "darkred")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0.5)) +
    labs(y = "AUC", 
         title = "Growth of strains in TSB",
         colour = "AMPO on plate")
```

### Minimal media

```{r, echo=F, warning=F, message=F, fig.width=20, fig.height=15, include=F}
results_AUC %>%  
  filter(Media %in% "MM") %>% 
  filter(Chem %in% c("DMSO", "MBOA")) %>% 
  ggplot(aes(x= interaction(as.factor(Chem)), y= AUC_raw)) +
  geom_boxplot(aes(fill = AMPO_on_MBOA_plates), position=position_dodge()) + 
  geom_jitter(width = 0.1) + 
  ggpubr::stat_compare_means(label = "p.signif", method = "t.test", ref.group = "DMSO", label.y.npc = 0.95) +
  facet_wrap(~Strain, scales = "free_y") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 1.1)))+
  scale_fill_manual(values = c("cornflowerblue", "darkred")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0.5)) +
    labs(y = "AUC", 
         title = "Growth of strains in TSB",
         colour = "AMPO on plate")
```

```{r, echo=F, warning=F, message=F, include=FALSE}
# results %>% filter(Media %in% "MM") %>% filter(Strain %in% "LMB2") %>% dplyr::select(Well, Treat, Replicate, Media, Chem, Run, Well_Plate,  Strain, Plate, Rep, Nr, Block) %>% unique()
```

## Plotting growth curves of minimal media
3 replicates (Rep) of each strain per plate
2 replicates (Replicate) per plate (twice the identical plate per run)
2 identical runs

```{r, echo=F, warning=F, message=F, fig.width=20, fig.height=15}
growth_curves_MBOA_Micros_AMPO <- results %>% 
  filter(AMPO_on_MBOA_plates %in% "yes") %>% 
  filter(Media %in% "MM") %>% 
  filter(!Strain %in% c("no", "NBC")) %>% 
  mutate(Chem_Media = interaction(Chem, Media, sep = "_")) %>% 
  dplyr::arrange(desc(AMPO_on_MBOA_plates)) %>% 
  ggplot(aes(x=hrs, y= Density_increase, color = Chem_Media)) +
  geom_line(aes(linetype = interaction(Replicate, Run, as.factor(Block))), stat = "identity", show.legend = TRUE) +
  facet_wrap(~reorder(Strain, desc(AMPO_on_MBOA_plates)), scales = "fixed") +
  scale_color_manual(values = c("#bebebe","#408a32","#555ccb"),
                     labels = c("DMSO", "Glucose 30 000 μM", "MBOA 500 μM"))+
  scale_linetype_manual("Replicate", values=c(1:12), labels=c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))+
  ylim(0,1.3)+
  theme_bw() +
  labs(x = "Time [hours]",
        y = "Bacterial growth [OD600]", 
        title = "Bacterial growth with MBOA as sole carbon source",
        subtitle = "AMPO-forming Microbactera",
        linetype = "Replicate",
        color = "Chemical")

growth_curves_MBOA_Micros_AMPO
```

```{r, echo=F, warning=F, message=F, fig.width=20, fig.height=15}
growth_curves_MBOA_Micros_AMPO_pos <- results %>% 
  # filter(AMPO_on_MBOA_plates %in% "yes") %>% 
  filter(Media %in% "MM") %>% 
  filter(!Strain %in% c("no", "NBC")) %>% 
  filter(Chem %in% c("DMSO", "MBOA")) %>%
  filter(Density_increase > 0) %>% 
  mutate(Chem_Media = interaction(Chem, Media, sep = "_")) %>% 
  dplyr::arrange(desc(AMPO_on_MBOA_plates)) %>% 
  ggplot(aes(x=hrs, y= Density_increase, color = Chem_Media)) +
  geom_line(aes(linetype = interaction(Replicate, Run, as.factor(Block))), stat = "identity", show.legend = TRUE) +
  facet_wrap(~reorder(Strain, desc(AMPO_on_MBOA_plates)), scales = "fixed") +
  scale_color_manual(values = c("#bebebe","#555ccb"),
                     labels = c("DMSO", "MBOA 500 μM"))+
  scale_linetype_manual("Replicate", values=c(1:12), labels=c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))+
  # ylim(0,1.3)+
  theme_bw() +
  labs(x = "Time [hours]",
        y = "Bacterial growth [OD600]", 
        title = "Bacterial growth with MBOA as sole carbon source",
        subtitle = "AMPO-forming Microbactera",
        linetype = "Replicate",
        color = "Chemical")

growth_curves_MBOA_Micros_AMPO_pos
```

```{r, echo=F, warning=F, message=F, fig.width=20, fig.height=15}
growth_curves_MBOA_Micros_no_AMPO <- results %>% 
  filter(AMPO_on_MBOA_plates %in% "no") %>% 
  filter(Media %in% "MM") %>% 
  filter(!Strain %in% c("no", "NBC")) %>% 
  mutate(Chem_Media = interaction(Chem, Media, sep = "_")) %>% 
  dplyr::arrange(desc(AMPO_on_MBOA_plates)) %>% 
  ggplot(aes(x=hrs, y= Density_increase, color = Chem_Media)) +
  geom_line(aes(linetype = interaction(Replicate, Run, as.factor(Block))), stat = "identity", show.legend = TRUE) +
  facet_wrap(~reorder(Strain, desc(AMPO_on_MBOA_plates)), scales = "fixed") +
  scale_color_manual(values = c("#bebebe","#408a32","#555ccb"),
                     labels = c("DMSO", "Glucose 30 000 μM", "MBOA 500 μM"))+
  scale_linetype_manual("Replicate", values=c(1:12), labels=c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))+
  ylim(0,1.3)+
  theme_bw() +
  labs(x = "Time [hours]",
        y = "Bacterial growth [OD600]", 
        title = "Bacterial growth with MBOA as sole carbon source",
        subtitle = "non-AMPO-forming Microbactera",
        linetype = "Replicate",
        color = "Chemical")

growth_curves_MBOA_Micros_no_AMPO
```


## Phylogenetic tree

```{r, include=FALSE}
results_AUC_mean_MM <- results_AUC %>% filter(Media %in% "MM") %>% group_by(Strain, Chem) %>%
  dplyr::summarise(mean = mean(AUC), n = n()) %>% 
  dplyr::rename(AUC_mean = mean,
                Replicates = n)

results_AUC_mean_MM_pos_sum <- results_AUC %>% 
  filter(Media %in% "MM") %>% 
  group_by(Strain, Chem) %>%
  dplyr::summarise(mean = mean(AUC_pos), n = n()) %>% 
  dplyr::rename(AUC_mean_pos = mean, Replicates = n) %>% 
  filter(Chem %in% "MBOA") 

results_AUC_mean_MM$Strain <- gsub("NBC", "KHB019", results_AUC_mean_MM$Strain)

results_AUC_mean_MM %<>% mutate(AUC_mean = as.numeric(AUC_mean))
# results_AUC_mean_MBOA <- results_AUC_mean %>% filter(Chem %in% "MBOA") %>% dplyr::select(-Chem)
```

```{r, include=FALSE}
results_AUC_mean_MM_MBOA <- results_AUC_mean_MM %>% filter(Chem %in% "MBOA") 
results_AUC_mean_MM_MBOA_pos <- results_AUC_mean_MM_MBOA
results_AUC_mean_MM_MBOA_pos$AUC_mean <- ifelse(results_AUC_mean_MM_MBOA_pos$AUC_mean < 0, 0, results_AUC_mean_MM_MBOA_pos$AUC_mean)
```

Fig 8c: Growth of strains in minimal medium with 500 uM MBOA. 

```{r, echo=F, warning=F, message=F, include=FALSE}
results_AUC %>% filter(Media %in% "MM") %>% ggplot(aes(x = Strain, y = AUC_raw)) +
  geom_bar(aes(fill = AMPO_on_MBOA_plates), stat = "summary") +
  scale_fill_manual(values = c("cornflowerblue", "darkred"))+
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) 
```

```{r, echo=F, warning=F, message=F, include=FALSE}
results_AUC %>% filter(Media %in% "MM") %>% 
  ggplot(aes(x = reorder(Strain, AMPO_on_MBOA_plates), y = AUC_pos)) +
  geom_bar(aes(fill = AMPO_on_MBOA_plates), stat = "summary") +
  scale_fill_manual(values = c("cornflowerblue", "darkred")) +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) 
```


*Binary interpretation of bacterial growth:*  
For binary interpretation (Bacterium X uses MBOA as C-source: YES or NO) of bacterial growth (quantitatively shown in Fig. S8c), we compared the 
AUC[MBOA] vs. AUC[DMSO]. The binary data is then reported in Fig. 3 ("C-Source"). We defined a cutoff of >1.5x for the AUC[MBOA]-AUC[DMSO] for bacterial growth on MBOA as sole carbon source. 

```{r, echo=F, warning=F, message=F, include=FALSE}
results_AUC_dif_MBOA_DMSO <- results_AUC %>% as.data.frame() %>% filter(Media %in% "MM") %>% 
  select(Strain, Well, Chem, AUC, AMPO_on_MBOA_plates, Plate, Replicate, Run) %>% 
  pivot_wider(values_from = AUC, names_from = Chem) %>% 
  mutate(Dif_MBOA_DMSO = MBOA-DMSO)
```

```{r, echo=F, warning=F, message=F}
results_AUC_dif_MBOA_DMSO %>% filter(!Dif_MBOA_DMSO %in% NA) %>% 
  ggplot(aes(x = Strain, y = Dif_MBOA_DMSO)) +
  geom_bar(aes(fill = AMPO_on_MBOA_plates), stat = "summary") +
  scale_fill_manual(values = c("cornflowerblue", "darkred")) +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) 

```

```{r, echo=F, warning=F, message=F, include=FALSE}
results_AUC_Dif_MBOA_DMSO_mean <- results_AUC_dif_MBOA_DMSO %>% 
  filter(!Dif_MBOA_DMSO %in% NA) %>% 
  group_by(Strain) %>%
  dplyr::summarise(mean = mean(Dif_MBOA_DMSO)) %>% 
  dplyr::rename(Dif_MBOA_DMSO_mean = mean)
```

```{r, echo=F, warning=F, message=F, include=FALSE}
results_AUC_Dif_MBOA_DMSO_mean %<>% mutate(Growth_MM_binary = case_when(Dif_MBOA_DMSO_mean > 1.5 ~ "growth", 
                                                                       Dif_MBOA_DMSO_mean < 1.5 ~ "no_growth"))
```

Figure S8c

```{r, echo=F, warning=F, message=F}
Fig.S8c.MBOA_Micro_carbon <- ggtree::gheatmap(p_host_simple,
         # results_AUC_mean_MM_pos_sum %>%
         # dplyr::select(Strain, AUC_mean_pos) %>% 
         results_AUC_Dif_MBOA_DMSO_mean %>% 
         dplyr::select(Strain, Dif_MBOA_DMSO_mean) %>% 
         as.data.frame() %>% column_to_rownames("Strain") ,
         offset=0.25,
         width=0.1,
         colnames=FALSE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_gradient2(low = "black", mid = "khaki", high ="darkred", midpoint = 1) +
  labs(x = "",
       y = "", 
       shape = "Host plant",
       color = "AMPO phenotype",
       fill = "Dif. AUC MBOA-DMSO")

Fig.S8c.MBOA_Micro_carbon

ggsave(plot = Fig.S8c.MBOA_Micro_carbon, filename = "FigS8c_Micro_carbon_source_tree.svg",width = 25, height = 15, dpi = 300, scale = 1, units = "cm")
```

# Combined tree

```{r, echo=F, warning=F, message=F, include=FALSE}
met %>% filter(Compound %in% c("MBOA", "HMPAA", "AMPO")) %>% 
  mutate(Compound = factor(Compound, levels = c("MBOA", "HMPAA", "AMPO"))) %>% 
  ggplot(aes(x= Strain, y = uMol)) + 
  geom_bar(stat = "summary") + 
  facet_grid(Treatment~Compound) +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) +
  labs(title = "MBOA metabolites")
```

```{r, echo=F, warning=F, message=F, include=F}
#HMPAA 
met %>% filter(Compound %in% c("HMPAA")) %>%  dplyr::summarize(conc_max = max(uMol)) 
met %>% filter(Compound %in% c("HMPAA")) %>%  filter(Treatment %in% "DMSO") %>% dplyr::summarize(conc_max = max(uMol)) 

met %>% filter(Compound %in% c("HMPAA")) %>%  dplyr::summarize(conc_min = min(uMol))
met %>% filter(Compound %in% c("HMPAA")) %>% filter(uMol > 0) %>% dplyr::summarize(conc_min = min(uMol))
met %>% filter(Compound %in% c("HMPAA")) %>% filter(uMol > 5) %>% dplyr::select(Strain, Treatment)

met %>% filter(Compound %in% c("AMPO")) %>% filter(uMol > 5) %>% dplyr::select(Strain, Treatment)
met %>% filter(Compound %in% c("AMPO")) %>% filter(uMol < 5) %>% dplyr::select(Strain, Treatment)
```


```{r, echo=F, warning=F, message=F, include=F}
metDA %>% 
  filter(Treatment %in% c("G", "D")) %>% 
  filter(Compound %in% c("DIMBOA-Glc", "MBOA", "AMPO")) %>% 
  mutate(Compound = factor(Compound, levels = c("DIMBOA-Glc", "MBOA", "AMPO"))) %>% 
  ggplot(aes(x= Strain, y = uMol)) + 
  geom_bar(stat = "summary") + 
  facet_grid(Treatment~Compound) +
  theme(axis.text.x=element_text(angle = -90, hjust = 0, vjust = 0 )) +
  labs(title = "DIMBOA-Glc metabolites")
```

Thresholds for qualitative analysis
- weak MBOA-degraders: >30% of MBOA degraded compared to the control)
- strong MBOA-degraders: (>90% of MBOA degraded, Fig. 3a). 
strong AMPO-formers (100 - 10 %) 
weak AMPO-formers’, <10% of max. AMPO forming strain)
(0.01 µM, limit of detection).


```{r, echo=F, warning=F, message=F, include=F}
met_MBOA_NBC <- met %>% filter(Strain %in% "NBC") %>% filter(Treatment %in% "MBOA") %>% filter(Compound %in% "MBOA")
met_MBOA_NBC$uMol
# 343.0771 uMol

# 70% 
MBOA_treshold_weak <- (met_MBOA_NBC$uMol / 100) * 70 
MBOA_treshold_weak
# 240.154

# 90 %
MBOA_treshold_strong <- (met_MBOA_NBC$uMol / 100) * 10
MBOA_treshold_strong
# 34.30771

# 0.1 %
MBOA_treshold_LOD <- (met_MBOA_NBC$uMol / 100) * 1
MBOA_treshold_LOD
```

```{r, echo=F, warning=F, message=F, include=F}
met_AMPO_max <- met %>% filter(Treatment %in% "MBOA") %>% filter(Compound %in% c("AMPO")) %>%  dplyr::summarize(conc_max = max(uMol)) 
met_AMPO_max$conc_max
# 471.5786

# 10% of max
AMPO_treshold_strong <- (met_AMPO_max$conc_max / 100) * 10
AMPO_treshold_strong
# 47.15786

# 0.1 % of max
AMPO_treshold_weak <- (met_AMPO_max$conc_max / 100) * 0.1
AMPO_treshold_weak
# 0.4715786
```

```{r, echo=F, warning=F, message=F, include=F}
met_HMPAA_max <- met %>% filter(Treatment %in% "MBOA") %>% filter(Compound %in% c("HMPAA")) %>%  dplyr::summarize(conc_max = max(uMol)) 
met_HMPAA_max$conc_max
# 44.06674

# 10% of max
HMPAA_treshold_strong <- (met_HMPAA_max$conc_max / 100) * 10
HMPAA_treshold_strong
# 4.406674

# 0.1 % of max
HMPAA_treshold_weak <- (met_HMPAA_max$conc_max / 100) * 0.1
HMPAA_treshold_weak
# 0.04406674
```

```{r, echo=F, warning=F, message=F, include=F}
met_DIMBOAGlc_NBC <- metDA %>% filter(Strain %in% "NBC") %>% filter(Treatment %in% "G") %>% filter(Compound %in% "DIMBOA-Glc")
met_DIMBOAGlc_NBC$uMol
# 515.6157

# 70% 
DIMBOAGlc_treshold_weak <- (met_DIMBOAGlc_NBC$uMol / 100) * 70 
DIMBOAGlc_treshold_weak
# 360.931

# 90 %
DIMBOAGlc_treshold_strong <- (met_DIMBOAGlc_NBC$uMol / 100) * 10
DIMBOAGlc_treshold_strong
# 51.56157

# 0.1 %
DIMBOAGlc_treshold_LOD <- (met_DIMBOAGlc_NBC$uMol / 100) * 0.1
DIMBOAGlc_treshold_LOD

# 0.5156157
```

```{r, echo=F, warning=F, message=F, include=F}
metDA_MBOA_max <- met %>% filter(Treatment %in% "MBOA") %>% filter(Compound %in% c("MBOA")) %>%  dplyr::summarize(conc_max = max(uMol)) 
metDA_MBOA_max$conc_max
# 516.2433

# 10% of max
MBOA.DG_treshold_strong <- (metDA_MBOA_max$conc_max / 100) * 10
MBOA.DG_treshold_strong
# 51.62433

# 0.1 % of max
MBOA.DG_treshold_weak <- (metDA_MBOA_max$conc_max / 100) * 0.1
MBOA.DG_treshold_weak
# 0.5162433
```

```{r, echo=F, warning=F, message=F, include=F}
met %>% filter(Strain %in% "Leaf151") %>% filter(Compound %in% c("MBOA", "HMPAA", "AMPO")) %>% filter(Treatment %in% "MBOA")

met %<>% as.data.frame() %>% filter(!RT %in% "NA")
```

```{r, echo=F, warning=F, message=F, include=F}
met <- met %>% as.data.frame() %>% mutate(Compoud = as.factor(Compound))
met$Binary <- NA #add new empty col

# met$uMol[is.na(met$uMol)] = 0

met$Binary[met$Compoud=="HMPAA" & met$uMol<HMPAA_treshold_weak] <- "no" # no HMPAA detected
met$Binary[met$Compoud=="HMPAA" & met$uMol< HMPAA_treshold_strong & met$uMol > HMPAA_treshold_weak] <- "HMPAA_low" #HMPAA not degraded & "MBOA high"
met$Binary[met$Compoud=="HMPAA" & met$uMol> HMPAA_treshold_strong] <- "HMPAA_high" #MBOA not degraded & "MBOA high"

met$Binary[met$Compoud=="AMPO" & met$uMol< AMPO_treshold_weak] <- "no" # no AMPO detected
met$Binary[met$Compoud=="AMPO" & met$uMol<AMPO_treshold_strong & met$uMol>AMPO_treshold_weak+0.001] <- "AMPO_low" #AMPO not degraded & "MBOA high"
met$Binary[met$Compoud=="AMPO" & met$uMol>AMPO_treshold_strong] <- "AMPO_high" #AMPO not degraded & "MBOA high"

met$Binary[met$Compoud=="MBOA" & met$uMol<MBOA_treshold_strong] <- "no" # no MBOA detected
met$Binary[met$Compoud=="MBOA" & met$uMol<MBOA_treshold_weak & met$uMol>MBOA_treshold_strong] <- "MBOA_low" #MBOA not degraded & "MBOA high"
met$Binary[met$Compoud=="MBOA" & met$uMol>MBOA_treshold_weak] <- "MBOA_high" #MBOA not degraded & "MBOA high"

metDA <- metDA %>% filter(Treatment %in% "G") %>% as.data.frame() %>% mutate(Compoud = as.factor(Compound))
metDA$Binary <- NA #add new empty col

metDA$Binary[metDA$Compoud=="DIMBOA-Glc" & metDA$uMol<DIMBOAGlc_treshold_strong] <- "no" # no MBOA detected
metDA$Binary[metDA$Compoud=="DIMBOA-Glc" & metDA$uMol<DIMBOAGlc_treshold_weak & metDA$uMol>DIMBOAGlc_treshold_strong] <- "DIMBOA-Glc_low" #MBOA not degraded & "MBOA high"
metDA$Binary[metDA$Compoud=="DIMBOA-Glc" & metDA$uMol>DIMBOAGlc_treshold_weak] <- "DIMBOA-Glc_high" #MBOA not degraded & "MBOA high"

metDA$Binary[metDA$Compoud=="MBOA" & metDA$uMol<MBOA.DG_treshold_weak] <- "no" # no MBOA detected
metDA$Binary[metDA$Compoud=="MBOA" & metDA$uMol<MBOA.DG_treshold_strong & metDA$uMol>MBOA.DG_treshold_weak] <- "MBOA_low" #MBOA not degraded & "MBOA high"
metDA$Binary[metDA$Compoud=="MBOA" & metDA$uMol>MBOA.DG_treshold_strong] <- "MBOA_high" #MBOA not degraded & "MBOA high"
```

```{r, echo=F, warning=F, message=F, include=F}
results_AUC_mean_MM_MBOA_pos %>% ggplot(aes(x = Strain, y = AUC_mean)) + geom_bar(stat = "summary")
```


```{r, echo=F, warning=F, message=F, include=F}
results_AUC_mean_MM_MBOA %>% filter(Strain %in% c("LWH11", "Leaf320"))

results_AUC_mean_MM_MBOA_max <- results_AUC_mean_MM_MBOA %>% as.data.frame() %>%  dplyr::summarize(conc_max = max(AUC_mean)) 
results_AUC_mean_MM_MBOA_max$conc_max

# 10% of max
results_AUC_mean_MM_MBOA_max_treshold <- (results_AUC_mean_MM_MBOA_max$conc_max / 100) * 20

results_AUC_mean_MM_MBOA %<>% mutate(Growth_MM = case_when(AUC_mean < results_AUC_mean_MM_MBOA_max_treshold ~ "no growth",
                                                           AUC_mean > results_AUC_mean_MM_MBOA_max_treshold ~ "growth"))
```


1. Tree with AMPO Phenotype on plates

```{r, warning=F, echo=F, message=F, fig.width=15, fig.height=12}
tree_plate <- ggtree::gheatmap(p_host_simple + new_scale_fill(),
         metadata_tree %>% 
         dplyr::select(Strain, AMPO_on_MBOA_plates) %>% 
           dplyr::rename('Plate assay' = AMPO_on_MBOA_plates) %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain"),
         offset=0.2,
         width=0.1,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("lightgoldenrod", "darkred")) +
  labs(fill = "AMPO phenotype")

# tree_plate
```

2. Tree as above with MBOA metabolites

```{r, echo=F, warning=F, message=F, fig.width=15, fig.height=12}
tree_plate_MBOA <- ggtree::gheatmap(tree_plate + new_scale_fill(),
         met %>% filter(Compound %in% c("MBOA", "HMPAA", "AMPO")) %>% 
         filter(Treatment %in% "MBOA") %>% 
         mutate(Compound = factor(Compound, levels = c("MBOA", "AMPO", "HMPAA"))) %>% 
         dplyr::select(Strain, Compound, Binary) %>% unique() %>% 
         dplyr::select(Strain, Compound, Binary) %>% tidyr::spread(Compound, Binary) %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain") ,
         offset=0.25,
         width=0.3,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c( "darkred", "white", "gold2", "white", "darkblue", "white"), na.value = "white") +
  labs(colour = "AMPO phenotype",
       shape = "Host plant",
       fill = "Concentration") 

# tree_plate_MBOA
```

3. Tree as above with DIMBOA-Glc metabolites 

```{r, echo=F, warning=F, message=F, fig.width=15, fig.height=12}
tree_plate_MBOA_DMG <- ggtree::gheatmap(tree_plate_MBOA + new_scale_fill(),
         metDA %>% filter(Treatment %in% c("G")) %>% 
           filter(Compound %in% c("DIMBOA-Glc", "MBOA")) %>% 
           filter(!uMol %in% 0.00000000) %>%
           filter(!uMol > 600) %>%
           mutate(Compound = factor(Compound, levels = c("DIMBOA-Glc", "MBOA"))) %>% 
         dplyr::select(Strain, Compound, Binary) %>% tidyr::spread(Compound, Binary) %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain") ,
         offset=0.37,
         width=0.2,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("darkorange2",  "orange", "darkblue", "#555CCB", "white"), na.value = "white")+
  labs(#x = "Benzoxazinoid",
       #y = "Relative AUC", 
       colour = "AMPO phenotype",
       shape = "Host plant",
       fill = "Concentration") 

# tree_plate_MBOA_DMG
```

4. Tree as above with carbon source data

```{r, echo=F, warning=F, message=F, fig.width=15, fig.height=12}
tree_plate_MBOA_DMG_carbon <- ggtree::gheatmap(tree_plate_MBOA_DMG + new_scale_fill(),
                                        # results_AUC_mean_MM_MBOA %>%
         results_AUC_Dif_MBOA_DMSO_mean %>% 
         dplyr::select(Strain, Growth_MM_binary) %>% 
         # dplyr::select(Strain, Growth_MM) %>% 
         # dplyr::rename('C-source' = Growth_MM) %>% 
         dplyr::rename('C-source' = Growth_MM_binary) %>% 
         as.data.frame() %>% column_to_rownames("Strain") ,
         offset=0.46,
         width=0.1,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         # colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("khaki4",  "white"))+
  labs(x = "",
       y = "", 
       shape = "Host plant",
       color = "AMPO phenotype",
       fill = "AUC")

# tree_plate_MBOA_DMG_carbon
```

5. Tree as above with orthogroups

Figure 3
note: phenotype of KHB019 on plate was adjusted in Illustrator to no AMPO.

```{r, echo=F, warning=F, message=F, fig.width=15, fig.height=12}
tree_plate_MBOA_DMG_carbon_OG <- ggtree::gheatmap(tree_plate_MBOA_DMG_carbon + new_scale_fill(),
         met_OG_selected %>% 
         distinct %>% 
         as.data.frame() %>% column_to_rownames("Strain"),
         offset=0.52,
         width=0.5,
         colnames=TRUE,
         font.size = 3,
         colnames_position = "top",
         colnames_offset_y = 5,
         colnames_angle = 90,
         colnames_level = NULL,
         hjust =1) +
  scale_fill_manual(values = c("gray100", "gray80", "gray60", "gray40", "gray20", "white")) +
  labs(fill = "Orthogroup copies")

tree_plate_MBOA_DMG_carbon_OG

ggsave(plot = tree_plate_MBOA_DMG_carbon_OG, filename = "Fig3_Tree_Microbacteria.svg", width = 26, height = 20, dpi = 300, scale = 1, units = "cm")
```

```{r, warning=F, echo=F, message=F}
sessionInfo()
```

